### 2020-01-03 sums_data_cleaning.Rmd
Clean SUMs data

---
title: "KidsAIR initial SUMs data work"
author: "Ethan Walker"
date: "Started 23 Dec 2019, Updated 3 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

# Add linked IDs to WMT SUMs data logs
```{r}
wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log.csv") %>%
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  select(home_winter_id, ibid, date_log, time_log)

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  filter(area == "WMT") %>% 
  select(home_winter_id, winter_id, home, home_id)

wmt_sums_log <- wmt_sums_data_log %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
  arrange(home, date_log) %>% 
  mutate(date_log = mdy(date_log))

write_rds(wmt_sums_log, "Output/wmt_sums_data_log.rds")
write_csv(wmt_sums_log, "Output/wmt_sums_data_log.csv")

wmt_sums_data_log_updated <- read_csv("Input/WMT/wmt_sums_data_log_updated.csv") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         notes = `REPLACE(REPLACE(Notes`) %>% 
  select(home_winter_id, date, time, alarms, file_name, notes)

wmt_sums_log <- wmt_sums_data_log_updated %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, home, home_id, date, time, alarms, file_name, notes) %>% 
  arrange(home, date)

write_csv(wmt_sums_log, "Output/wmt_sums_data_log_updated.csv")
```

# Add linked IDs to NN SUMs data logs
```{r}
nn_sums_data_log <- read_xlsx("Input/NN/nn_sums_data_log.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  select(home_winter_id, ibid, date_log, time_log)

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  filter(area == "NN") %>% 
  select(home_winter_id, winter_id, home, home_id)

nn_sums_log <- nn_sums_data_log %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
  arrange(home, date_log) %>% 
  separate(time_log, c("trash", "time_log"), sep = " ") %>% 
  select(-trash)

write_rds(nn_sums_log, "Output/nn_sums_data_log.rds")
write_csv(nn_sums_log, "Output/nn_sums_data_log.csv")

nn_sums_data_log_updated <- read_xlsx("Input/NN/nn_sums_data_log_updated.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         notes = IgnoreReason) %>% 
  select(home_winter_id, date, time, alarms, file_name, notes)

nn_sums_log <- nn_sums_data_log_updated %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, home, home_id, date, time, alarms, file_name, notes) %>% 
  arrange(home, date)

write_csv(nn_sums_log, "Output/nn_sums_data_log_updated.csv")
```

# Add linked IDs to AK SUMs data logs
```{r}
ak_sums_data_log <- read_xlsx("Input/AK/ak_sums_data_log.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  separate(time_log, c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log)

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  filter(area == "AK") %>% 
  select(home_winter_id, winter_id, home, home_id)

ak_sums_log <- ak_sums_data_log %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
  arrange(home, date_log)

write_csv(ak_sums_log, "Output/ak_sums_data_log.csv")
write_rds(ak_sums_log, "Output/ak_sums_data_log.rds")

ak_sums_data_log_updated <- read_xlsx("Input/AK/ak_sums_data_log_updated.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         notes = IgnoreReason) %>% 
  select(home_winter_id, date, time, alarms, file_name, notes)

ak_sums_log <- ak_sums_data_log_updated %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, home, home_id, date, time, alarms, file_name, notes) %>% 
  arrange(home, date)

write_csv(ak_sums_log, "Output/ak_sums_data_log_updated.csv")
```

# Initial load, format, and save KidsAIR WMT SUMs data
## Files here are in several formats: csv (3 different formats) and txt (5 formats)
## I had to do some manual cleaning of the files to get them named
## consistently and in reasonably similar formats to be able to load
## into R in batches, as seen below. The files I upload in the code below
## will also be uploaded to Box in a folder with my name.
## Box folder directory: 
```{r}
##### Read in sums csv files (from WMT) #####

# List files
list_csv_files <- list.files("Input/WMT/sums_csv_new")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/WMT/sums_csv_new")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 1,
                            col_names = c("line", "date", "time", "temp_c", "trash"),
                            col_types = cols(line = col_double(),
                                             date = col_date(format = ""),
                                             time = col_time(format = ""),
                                             temp_c = col_character(),
                                             trash = col_character()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_sums_csv <- initial_data %>% 
  mutate(area = "WMT",
         home2 = "WMA") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c_sums = temp_c) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_csv_files_2 <- list.files("Input/WMT/sums_csv_old/format_1")

# Set working directory and load files in list; extract file name and add as column
## run next 7 lines together
setwd("Input/WMT/sums_csv_old/format_1")
initial_data = tibble(file_list = list_csv_files_2) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 22,
                            col_names = c("datetime_sums", "temp_c_sums"))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_sums_csv_2 <- initial_data %>% 
  mutate(area = "WMT",
         home2 = "WMA",
         datetime_sums = ymd_hms(datetime_sums)) %>% 
  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(date_sums = ymd(date_sums)) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_csv_files_3 <- list.files("Input/WMT/sums_csv_old/format_2")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/WMT/sums_csv_old/format_2")
initial_data = tibble(file_list = list_csv_files_3) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 2,
                            col_names = c("line", "date", "time", "temp_f", "trash"),
                            col_types = cols(line = col_double(),
                                             date = col_character(),
                                             time = col_time(format = ""),
                                             temp_f = col_double(),
                                             trash = col_character()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_sums_csv_3 <- initial_data %>% 
  mutate(area = "WMT",
         home2 = "WMA") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = mdy_hms(datetime),
         date_sums = mdy(date),
         time_sums = time,
         temp_c_sums = (temp_f-32)/1.8) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

##### Read in sums xlsx files (from WMT) #####

# List files
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip16")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/sums_xlsx_files/header_skip16")
initial_data = tibble(file_list = list_xlsx_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_xlsx, skip = 16)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_sums_xlsx <- initial_data %>% 
  mutate(datetime = ymd_hms(`Log data: date/time`),
         temp_c = as.numeric(`Temperature °C`),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c_sums = temp_c,
         home = as.factor(home)) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip21")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/sums_xlsx_files/header_skip21")
initial_data = tibble(file_list = list_xlsx_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_xlsx, skip = 21)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_sums_xlsx2 <- initial_data %>% 
  mutate(datetime = ymd_hms(`Log data: date/time`),
         temp_c = as.numeric(`Temperature °C`),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c_sums = temp_c,
         home = as.factor(home)) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_2col_c")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/sums_xlsx_files/no_header_2col_c")
initial_data = tibble(file_list = list_xlsx_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_xlsx)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_sums_xlsx3 <- initial_data %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c_sums = temp_c,
         home = as.factor(home)) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_c")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/sums_xlsx_files/no_header_4col_c")
initial_data = tibble(file_list = list_xlsx_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_xlsx)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_sums_xlsx4 <- initial_data %>% 
  mutate(temp_c = as.numeric(temp_c),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("time", c("trash", "time"), sep = " ") %>%
  unite("time", c("time", "am_pm"), sep = " ") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = ymd_hms(datetime),
         date_sums = ymd(date),
         time_sums = time,
         temp_c_sums = temp_c,
         home = as.factor(home)) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_f")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/sums_xlsx_files/no_header_4col_f")
initial_data = tibble(file_list = list_xlsx_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_xlsx)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_sums_xlsx5 <- initial_data %>% 
  mutate(temp_f = as.numeric(temp_f),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("time", c("trash", "time"), sep = " ") %>%
  unite("time", c("time", "am_pm"), sep = " ") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = ymd_hms(datetime),
         date_sums = ymd(date),
         time_sums = time,
         temp_c_sums = (temp_f-32)/1.8,
         home = as.factor(home)) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# Bind all WMT SUMs data together
wmt_sums_raw <- rbind(wmt_sums_csv, wmt_sums_csv_2, wmt_sums_csv_3,
                      wmt_sums_xlsx, wmt_sums_xlsx2, wmt_sums_xlsx3,
                      wmt_sums_xlsx4, wmt_sums_xlsx5) %>% 
  select(-datetime_sums) %>% 
  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  filter(!is.na(date_sums)) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

write_rds(wmt_sums_raw, "Output/wmt_sums_raw.rds")
```

# Initial load, format, and save KidsAIR NN SUMs data
```{r}
##### Read in sums csv files (from NN) #####

# List files
list_csv_files <- list.files("Input/NN/sums_csv_new")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/NN/sums_csv_new")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv,
                            col_names = c("line", "date", "time", "temp_c"),
                            col_types = cols(line = col_double(),
                                             date = col_date(format = ""),
                                             time = col_time(format = ""),
                                             temp_c = col_character()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
nn_sums_csv <- initial_data %>% 
  mutate(area = "NN") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c_sums = temp_c) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_csv_files_2 <- list.files("Input/NN/sums_csv_old/2col-c")

# Set working directory and load files in list; extract file name and add as column
## run next 7 lines together
setwd("Input/NN/sums_csv_old/2col-c")
initial_data = tibble(file_list = list_csv_files_2) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
nn_sums_csv_2 <- initial_data %>% 
  mutate(area = "NN",
         datetime_sums = mdy_hm(datetime)) %>% 
  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>%
  mutate(date_sums = ymd(date_sums),
         temp_c_sums = as.numeric(temp_c)) %>%
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_csv_files_3 <- list.files("Input/NN/sums_csv_old/3col-f")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/NN/sums_csv_old/3col-f")
initial_data = tibble(file_list = list_csv_files_3) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 1,
                            col_names = c("date", "time", "temp_f"),
                            col_types = cols(date = col_character(),
                                             time = col_time(format = ""),
                                             temp_f = col_double()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
nn_sums_csv_3 <- initial_data %>% 
  mutate(area = "NN") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_sums = mdy_hms(datetime),
         date_sums = mdy(date),
         time_sums = time,
         temp_c_sums = (temp_f-32)/1.8) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# Bind all WMT SUMs data together
nn_sums_raw <- rbind(nn_sums_csv, nn_sums_csv_2, nn_sums_csv_3) %>% 
  select(-datetime_sums) %>% 
  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  filter(!is.na(date_sums)) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

write_rds(nn_sums_raw, "Output/nn_sums_raw.rds")
```

# Initial load, format, and save KidsAIR AK SUMs data
```{r}
##### Read in sums csv files (from AK) #####

# List files
list_csv_files <- list.files("Input/AK/sums_csv_new")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/AK/sums_csv_new")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv,
                            col_names = c("line", "date", "time", "temp_c"),
                            col_types = cols(line = col_double(),
                                             date = col_date(format = ""),
                                             time = col_time(format = ""),
                                             temp_c = col_double()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
ak_sums_csv <- initial_data %>% 
  mutate(area = "AK") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>%
  mutate(datetime_sums = ymd_hms(datetime),
         date_sums = date,
         time_sums = time,
         temp_c_sums = temp_c) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_csv_files_2 <- list.files("Input/AK/sums_csv_old")

# Set working directory and load files in list; extract file name and add as column
## run next 7 lines together
setwd("Input/AK/sums_csv_old")
initial_data = tibble(file_list = list_csv_files_2) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 1,
                            col_names = c("line", "date", "time", "temp_f", "trash"),
                            col_types = cols(line = col_double(),
                                             date = col_character(),
                                             time = col_time(format = ""),
                                             temp_f = col_double(),
                                             trash = col_character())))  %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
ak_sums_csv_2 <- initial_data %>% 
  mutate(area = "AK") %>% 
  mutate(date_sums = mdy(date),
         time_sums = time,
         temp_c_sums = (temp_f-32)/1.8) %>% 
  unite("datetime", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_sums = ymd_hms(datetime)) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

##### Read in sums xlsx files (from AK) #####

# List files
list_xlsx_files <- list.files("Input/AK/sums_xlsx_temp_c")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/AK/sums_xlsx_temp_c")
initial_data = tibble(file_list = list_xlsx_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_xlsx,
                            col_names = c("date", "time", "am_pm", 
                                          "trash", "temp_c"))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
ak_sums_xlsx <- initial_data %>% 
  mutate(area = "AK") %>% 
  separate("time", c("trash", "time"), sep = " ", remove = FALSE) %>% 
  unite("time", c("time", "am_pm"), sep = " ") %>% 
  mutate(date_sums = ymd(date),
         time_sums = time,
         temp_c_sums = as.numeric(temp_c),
         home = as.factor(home)) %>% 
  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_xlsx_files2 <- list.files("Input/AK/sums_xlsx_temp_f")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/AK/sums_xlsx_temp_f")
initial_data = tibble(file_list = list_xlsx_files2) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_xlsx,
                            col_names = c("date", "time", "am_pm", 
                                          "trash", "temp_f"))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
ak_sums_xlsx2 <- initial_data %>% 
  mutate(area = "AK") %>% 
  separate("time", c("trash", "time"), sep = " ", remove = FALSE) %>% 
  unite("time", c("time", "am_pm"), sep = " ") %>% 
  mutate(date_sums = ymd(date),
         time_sums = time,
         temp_c_sums = as.numeric((temp_f-32)/1.8),
         home = as.factor(home)) %>% 
  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# Bind all AK SUMs data together
ak_sums_raw <- rbind(ak_sums_csv, ak_sums_csv_2, 
                     ak_sums_xlsx, ak_sums_xlsx2) %>% 
  select(-datetime_sums) %>% 
  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  filter(!is.na(date_sums)) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# write_rds(ak_sums_raw, "Output/ak_sums_raw.rds")
```

# Bind all SUMs files together
```{r}
wmt_sums_raw <- read_rds("Output/wmt_sums_raw.rds")

nn_sums_raw <- read_rds("Output/nn_sums_raw.rds")

ak_sums_raw <- read_rds("Output/ak_sums_raw.rds")

all_sums_raw <- rbind(wmt_sums_raw, nn_sums_raw, ak_sums_raw)

# write_rds(all_sums_raw, "Output/all_sums_raw.rds")
```

# Check sample start times vs log start times
```{r}
# Load all SUMs data
all_sums_raw <- read_rds("Output/all_sums_raw.rds") %>% 
  ungroup() %>% 
  arrange(area, home, datetime_sums)

# Load SUMs data logs
wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds")
nn_sums_log <- read_rds("Output/nn_sums_data_log.rds")
ak_sums_log <- read_rds("Output/ak_sums_data_log.rds")

# Combine SUMs logs from all areas
all_sums_log <- rbind(wmt_sums_log, nn_sums_log, ak_sums_log) %>% 
  mutate(winter_id = as.character(winter_id))

# Combine SUMs logs and data
## Get ready to compare start times from logs and data files
sums_time_comparison <- all_sums_raw %>%
  group_by(area, home) %>% 
  mutate(datetime_initial = first(datetime_sums)) %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums),
         datetime_initial = ymd_hms(datetime_initial),
         winter_id = if_else((datetime_sums - datetime_initial) > 20736000,
                             2, 1), 
         winter_id = as.character(winter_id)) %>% 
  ungroup() %>% 
  left_join(all_sums_log, by = c("home", "winter_id")) %>% 
  group_by(home, winter_id) %>% 
  unite(datetime_log, c("date_log", "time_log"), 
        sep = " ", remove = FALSE) %>% 
  mutate(datetime_log = ymd_hms(datetime_log),
         datetime_initial = first(datetime_sums),
         time_diff_sums = datetime_initial - datetime_log) %>% 
  distinct(time_diff_sums, .keep_all = TRUE) %>%
  ungroup() %>% 
  arrange(home, winter_id)

write_csv(sums_time_comparison, "Output/sums_time_comparison.csv")
write_rds(sums_time_comparison, "Output/sums_time_comparison.rds")
```