### 2020-01-10 dustrak_data_cleaning.Rmd
SUMs cleaning updates. Add extensive notes to PM and SUMs code.

---
title: "KidsAIR initial PM data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 10 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

# Initial load, format, and save KidsAIR DustTrak (PM) data
## Files were downloaded from Box: KidsAIR > 2019-Nov-Data > WMT > uploads > Dusttrack
## Files were not named consistently, so I renamed them by their home ID number.
## Files also had several different types and formats.
## Manual cleaning of some files had to be done to get them in 
## consistent formats and file types to work with in R.
## Unfortunately, this process had to be manual due to the large number of files
## and inconsistent QC process across the various files.
## As a result, this process will be difficult to replicate if needed;
## The code below will highlight how the files were formatted and grouped together.
## Once files were in consistent formats, I load them to R in batches, as seen below.
## The following code cleans and combines all PM data
```{r}
##### Read in dusttrack .txt files (from WMT) #####

# List files
## Numerous files had this format with a 28 row header
## 3 columns: date, time, PM
list_txt_files <- list.files("Input/WMT/dusttrack_txt_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/dusttrack_txt_files")
initial_data = tibble(file_list = list_txt_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 28)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_pm_txt <- initial_data %>% 
  mutate(date = mdy(`MM/dd/yyyy`),
         time = `hh:mm:ss`,
         pm = as.numeric(`mg/m^3`),
         area = "WMT") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, datetime, date, time, pm, area)

##### Read in dusttrack .csv files (all areas) #####

### WMT ###
# List files
## Numerous files had this format with no header
## 3 columns: date, time, PM
list_csv_files <- list.files("Input/WMT/dusttrack_csv_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/dusttrack_csv_files")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_pm_csv <- initial_data %>% 
  mutate(date = mdy(Date),
         time = Time,
         pm = as.numeric(AEROSOL),
         area = "WMT") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  filter(!is.na(pm))

### NN ###
# List files
## Numerous files had this format with no header
## 3 columns: date, time, PM
list_csv_files <- list.files("Input/NN/dusttrack_csv_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/NN/dusttrack_csv_files")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
nn_pm_csv <- initial_data %>% 
  mutate(date = mdy(Date),
         time = Time,
         pm = as.numeric(AEROSOL),
         area = "NN") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  filter(!is.na(pm))

### AK ###
# List files
## Numerous files had this format with no header
## 3 columns: date, time, PM
list_csv_files <- list.files("Input/AK/dusttrack_csv_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/AK/dusttrack_csv_files")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
ak_pm_csv <- initial_data %>% 
  mutate(date = mdy(Date),
         time = Time,
         pm = as.numeric(AEROSOL),
         area = "AK") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  filter(!is.na(pm))

##### Read in other dusttrack .csv files (WMT and NN) #####

### WMT ### 
# List files
## 4 files had this format with a 19 row header
## 4 columns: elapsed time (start time extracted from header), PM, notes, notes
list_other_files <- list.files("Input/WMT/dusttrack_other_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/dusttrack_other_files")
initial_data = tibble(file_list = list_other_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_pm_other <- initial_data %>% 
  mutate(trash1 = `Instrument Name`,
         trash2 = `DustTrak II`) %>% 
  select(home_winter_id, trash1, trash2) %>% 
  group_by(home_winter_id) %>% 
  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash2, "NA"),
         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash1, "NA"),
         trash1 = as.numeric(trash1),
         trash1 = as.character(trash1),
         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
  filter(!is.na(trash5)) %>% 
  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
  spread(trash4, trash3) %>% 
  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
  mutate(datetime = lead(datetime)) %>% 
  replace_with_na(replace = list(datetime = "NA NA")) %>% 
  mutate(datetime2 = first(datetime),
         datetime = mdy_hms(datetime2),
         trash1 = as.numeric(trash1)) %>% 
  filter(!is.na(trash1)) %>% 
  mutate(datetime2 = datetime + seconds(trash1),
         datetime2 = ymd_hms(datetime2)) %>% 
  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
  mutate(pm = as.numeric(trash2),
         datetime = datetime2,
         area = "WMT") %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  ungroup()

### NN ### 
# List files
## Numerous files had this format with a 19 row header
## 4 columns: elapsed time (start time extracted from header), PM, notes, notes
list_other_files <- list.files("Input/NN/dusttrack_other_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/NN/dusttrack_other_files")
initial_data = tibble(file_list = list_other_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
nn_pm_other <- initial_data %>% 
  mutate(trash1 = `Instrument Name`,
         trash2 = `DustTrak II`) %>% 
  select(home_winter_id, trash1, trash2) %>% 
  filter(!is.na(trash1)) %>% 
  group_by(home_winter_id) %>% 
  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash1, "NA"),
         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash2, "NA"),
         trash1 = as.numeric(trash1),
         trash1 = as.character(trash1),
         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
  filter(!is.na(trash5)) %>% 
  spread(trash3, trash4) %>% 
  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
  mutate(datetime = lead(datetime)) %>% 
  replace_with_na(replace = list(datetime = "NA NA")) %>% 
  mutate(datetime2 = first(datetime),
         datetime = mdy_hms(datetime2),
         trash1 = as.numeric(trash1)) %>% 
  filter(!is.na(trash1)) %>% 
  mutate(datetime2 = datetime + seconds(trash1),
         datetime2 = ymd_hms(datetime2)) %>% 
  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
  mutate(pm = as.numeric(trash2),
         datetime = datetime2,
         area = "NN") %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  ungroup() %>% 
  mutate(home_winter_id = if_else(home_winter_id == "220001" | home_winter_id == "220002",
                                  "22", home_winter_id),
         home_winter_id = if_else(home_winter_id == "550001" | home_winter_id == "550002",
                                  "55", home_winter_id),
         home_winter_id = if_else(home_winter_id == "810001" | home_winter_id == "810002",
                                  "81", home_winter_id))

##### Bind all files from above #####
kids_pm <- rbind(wmt_pm_csv, wmt_pm_txt, wmt_pm_other,
                nn_pm_csv, nn_pm_other,
                ak_pm_csv)

# Save pm dataset as RDS and CSV
# write_rds(kids_pm, "Output/kids_pm_full_raw.rds")
# write_csv(kids_pm, "Output/kids_pm_full.csv")
```

# Load Dusttrak data and do initial clean-up
```{r}
# Load full, raw dataset compiled in chunk above
kids_pm_full <- read_rds("Output/kids_pm_full_raw.rds")

# Load and combine data sheets with QAQC comments, etc
ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv") %>% 
  mutate(area = "AK")
ak_dt <- read_xlsx("Input/AK/ak_dustrak.xlsx") %>% 
  mutate(area = "AK") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv") %>% 
  mutate(area = "NN")
nn_dt <- read_xlsx("Input/NN/nn_dustrak.xlsx") %>% 
  mutate(area = "NN") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv") %>% 
  mutate(area = "WMT")
wmt_dt <- read_xlsx("Input/WMT/wmt_dustrak.xlsx") %>% 
  mutate(area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)

dt <- rbind(ak_dt, nn_dt, wmt_dt) %>% 
  mutate(home_winter_id = as.character(homewinterid)) %>% 
  select(-homewinterid)

dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
  rename_all(tolower) %>% 
  filter(!is.na(homewinterid)) %>% 
  mutate(home_winter_id = as.character(homewinterid)) %>% 
  select(-homewinterid) %>% 
  left_join(dt, by = c("area", "home_winter_id"))

# Join dt_home with full dataset
kids_pm_joined <- kids_pm_full %>% 
  left_join(dt_home, by = c("area", "home_winter_id")) %>% 
  rename(treatment = condition)

##### Add new variables to Kids PM data #####
kids_pm_new <- kids_pm_joined %>%  
  #filter(area == "AK") %>% 
  # apply correction factor to pm and change to ug
  mutate(pm = pm/1.65,
         pm = pm*1000) %>% 
  mutate(home_winter_id = as.character(home_winter_id),
         area = as.factor(area),
         pm_date = date,
         pm_time = time,
         pm = as.numeric(pm),
         day_of_week = weekdays(date)) %>% 
  unite("pm_datetime", date:time, sep = " ") %>% 
  #select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
  arrange(area, home_winter_id, pm_date) %>% 
  group_by(area, home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm),
         # breaking sampling period into days
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  # calculate pm means for each sampling day
  group_by(area, home_winter_id, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm),
         pm_sampling_day = sampling_day) %>% 
  separate(starttime, c("trash1", "starttime"), sep = " ") %>% 
  separate(stoptime, c("trash2", "stoptime"), sep = " ") %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, 
         -datetime, -trash1, -trash2)

# Save pm dataset as RDS and CSV
# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
# write_csv(kids_pm_new, "Output/kids_pm_new.csv")
```

# Summarize by ID and Dusttrak ID to look for trends
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new.rds")

# summary by home_winter_id
summary_id <- kids_pm %>% 
  group_by(area, home_winter_id) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Act vs Exp Diff" = first(as.numeric(actual_vs_expected_diff)),
            "Act vs Exp" = first(actual_vs_expected),
            "Comments" = first(qaqccomments))
 write_csv(summary_id, "output/kids_summary_id.csv")

# summary by dusttrak id
summary_dtid <- kids_pm %>% 
  group_by(dtid) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
```

# Histogram for distribution of PM by individual ID
```{r}
# Load data
kids_pm <- read_rds("Output/kids_pm_new.rds")

histogram_id <- kids_pm %>% 
  filter(home_winter_id == "10" & area == "AK") %>% 
  ggplot() + 
    geom_histogram(aes(pm), bins = 50) +
    theme_classic()
histogram_id
```

# Function to look at PM data across time
```{r}
# Plot PM by datetime to look at trends, gaps, and peaks
pm_time_plot_function <- function(id, location) {
pm_time_plot <- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  #filter(pm < 30000) %>% 
  arrange(pm_datetime) %>% 
  mutate(pm_sampling_day = as.character(pm_sampling_day),
         day_of_week = factor(day_of_week,
                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
                                         "Thursday", "Friday", "Saturday"))) %>%
  ggplot() + 
    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
    theme_classic() +
    labs(title = id,
         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
         x = "Datetime",
         color = "Day of week") +
    theme(axis.title.y = element_text(size = 12,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_text(size = 12),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 12, color = "black"),
          axis.line.x = element_line(colour = "black", size = 1), 
          axis.line.y = element_line(colour = "black", size = 1), 
          axis.ticks = element_blank()) +
    scale_color_manual(values = jv_palette)

gaps <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  mutate(pm_datetime = ymd_hms(pm_datetime),
         datetime_lead = ymd_hms(lead(pm_datetime)),
         datetime_1 = seconds(pm_datetime),
         datetime_2 = seconds(datetime_lead),
         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
  arrange(desc(datetime_diff_minutes)) %>% 
  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
         pm_datetime, datetime_lead)

pm_desc <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  arrange(desc(pm)) %>% 
  select(pm_datetime, pm, pm_sampling_day)

pm_time_plot
}
```

# Run function from above to look at results
## Function run for each household to check for odd trends/gaps in the data
## Notes from this process on uploaded to Box:
### KidsAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new.rds")

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "129", location = "WMT")
head(gaps, 25)
head(pm_desc, 25)
tail(pm_desc, 25)

# filter data for a single home
check_data <- kids_pm %>% 
  filter(area == "WMT" & home_winter_id == "129") %>% 
  filter(pm < 0)
```

# Check sample start times
```{r}
# Load data
kids_pm <- read_rds("Output/kids_pm_new.rds")

# Checking difference between dustrak start datetime and field data log datetime
sample_times <- kids_pm %>% 
  #filter(area == "AK") %>% 
  group_by(area, home_winter_id) %>% 
  arrange(pm_datetime) %>% 
  unite(start_datetime, c("startdate", "starttime"), 
        sep = " ", remove = FALSE) %>% 
  mutate(first_datetime = first(pm_datetime),
         first_datetime = ymd_hms(first_datetime),
         start_datetime = ymd_hms(start_datetime),
         time_diff = first_datetime - start_datetime) %>% 
  distinct(time_diff, .keep_all = TRUE) %>% 
  select(area, home_winter_id, pm_datetime, first_datetime,
         start_datetime, time_diff, startdate, starttime) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id)
# write_csv(sample_times, "Output/pm_sample_times_comparison.csv")
```

# Loading and combining notes from time and PM logs
## Notes from the process of comparing actual vs expect start times 
## on uploaded to Box:
### KidsAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
```{r}
time_notes <- read_xlsx("Input/pm_sample_times_comparison.xlsx") %>% 
  select(area, home_winter_id, first_datetime, start_datetime,
         time_diff, next_steps_time, check_log_time)

pm_notes <- read_xlsx("Input/kids_summary_id.xlsx") %>% 
  select(area, home_winter_id, n, further_comments_pm, next_steps_pm, check_log_pm)

kids_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  mutate(home_winter_id = as.numeric(home_winter_id))

notes_joined <- pm_notes %>% 
  left_join(time_notes, by = c("area", "home_winter_id")) %>% 
  left_join(kids_ids, by = c("area", "home_winter_id"))

# write_csv(notes_joined, "Output/pm_notes_check_logs.csv")
```

# Fix dusttrak sample start times
## Notes from above were used to make the following changes in the data
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new.rds")

# Checking difference between dustrak start datetime and field data log datetime
kids_pm_new <- kids_pm %>% 
  #filter(area == "NN") %>% 
  group_by(area, home_winter_id) %>% 
  arrange(pm_datetime) %>% 
  unite(start_datetime, c("startdate", "starttime"), 
        sep = " ", remove = FALSE) %>% 
  mutate(first_datetime = first(pm_datetime),
         first_datetime = ymd_hms(first_datetime),
         start_datetime = ymd_hms(start_datetime),
         time_diff = first_datetime - start_datetime) %>% 
  #distinct(time_diff, .keep_all = TRUE) %>% 
  # make general time changes based on ranges
  mutate(pm_datetime = ymd_hms(pm_datetime),
         pm_datetime_new = if_else(time_diff > 1800 & time_diff < 5400, 
                                   pm_datetime - 3600, pm_datetime),
         pm_datetime_new = if_else(time_diff < -1800 & time_diff > -5400, 
                                   pm_datetime_new + 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff > 5400 & time_diff < 9000, 
                                   pm_datetime_new - 7200, pm_datetime_new),
         pm_datetime_new = if_else(time_diff < -5400 & time_diff > -9000, 
                                   pm_datetime_new + 7200, pm_datetime_new)) %>% 
  # make specific time changes
  mutate(pm_datetime_new = if_else(time_diff == 10903, # WMA457
                                   pm_datetime_new - 10800, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == (-82721), # WMA440
                                   pm_datetime_new + 82800, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 10069, # WMA459
                                   pm_datetime_new - 9900, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 9817, # WMA314
                                   pm_datetime_new - 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff < -31535900, # YKC06-103 
                                   pm_datetime_new + 31539519, pm_datetime_new),
         pm_datetime_new = if_else(time_diff < -31532200, # WMA305
                                   pm_datetime_new - 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 9120, # WMA228
                                   pm_datetime_new - 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == (-21497), # WMA130
                                   pm_datetime_new + 21497, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 684, # WMA237
                                   pm_datetime_new + 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 149793, # WMA238
                                   pm_datetime_new - 149793, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == -46227, # CH320
                                   pm_datetime_new + 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 13514, # TC360 
                                   pm_datetime_new - 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == -10696, # TC357 
                                   pm_datetime_new + 10800, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == -43427, # TC355 
                                   pm_datetime_new + 43200, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == -11388, # TC353 
                                   pm_datetime_new + 10800, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == -221, # TC351 
                                   pm_datetime_new + 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 4999, # CH214 
                                   pm_datetime_new + 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 4061, # WMA203 
                                   pm_datetime_new + 3600, pm_datetime_new)) %>% 
  mutate(first_datetime_new = first(pm_datetime_new),
         first_datetime_new = ymd_hms(first_datetime_new),
         start_datetime = ymd_hms(start_datetime),
         time_diff_new = first_datetime_new - start_datetime) %>%  
  #distinct(time_diff, .keep_all = TRUE) %>%
  ungroup() %>% 
  arrange(area, home_winter_id, pm_datetime_new)

# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
```

# Further cleaning:
## Remove specific homes from PM dataset
## Remove extremely high values
## Change negative values
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new.rds")

kids_pm_new <- kids_pm %>%
  # filtering out specific homes where PM data is not suitable for analysis
  # removes 119,520 obs
         # YKC06_105: only 12 hours of data
  mutate(do_not_use = if_else(area == "AK" & home_winter_id == "39", 1, 0),
         # YKC06_313: sample hourly instead of every minute
         do_not_use = if_else(area == "AK" & home_winter_id == "89", 1, do_not_use),
         # CH_302: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "127", 1, do_not_use),
         # CH_304: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "1330", 1, do_not_use),
         # CH_306: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "134", 1, do_not_use),
         # CH_307: only 8 datapoints
         do_not_use = if_else(area == "NN" & home_winter_id == "135", 1, do_not_use),
         # TC_357: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "1510", 1, do_not_use),
         # CH_322: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "161", 1, do_not_use),
         # CH_325: 6300 obs are negative, down to -12.12. Median is negative.
         do_not_use = if_else(area == "NN" & home_winter_id == "165", 1, do_not_use),
         # CH_400: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "2150", 1, do_not_use),
         # CH_403: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "2210", 1, do_not_use),
         # CH_108: Only 300 datapoints and no other files found
         do_not_use = if_else(area == "NN" & home_winter_id == "29", 1, do_not_use),
         # CH_207: Only 400 datapoints; no other files found
         do_not_use = if_else(area == "NN" & home_winter_id == "58", 1, do_not_use),
         # TC_230: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "620", 1, do_not_use),
         # CH_223: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "840", 1, do_not_use),
         # TC_242: 1 day worth of data with lots of gaps
         do_not_use = if_else(area == "NN" & home_winter_id == "86", 1, do_not_use),
         # WMA277: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "1910", 1, do_not_use),
         # WMA305: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2500", 1, do_not_use),
         # WMA306: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2510", 1, do_not_use),
         # WMA358: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2860", 1, do_not_use),
         # WMA448: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "3750", 1, do_not_use),
         # WMA200: 4700 negative values, down to -4.85. Median is negative
         do_not_use = if_else(area == "WMT" & home_winter_id == "152", 1, do_not_use),
         # WMA201: 5000 negative values, down to -10. Median negative. 17 hour gap 
         do_not_use = if_else(area == "WMT" & home_winter_id == "153", 1, do_not_use),
         # WMA603: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "5100", 
                              1, do_not_use)) %>% 
  filter(do_not_use != 1) %>% 
  # replace PM > 10,000 with NA (33 obs)
  mutate(pm = if_else(pm > 10000, 999999, pm)) %>% 
  replace_with_na(replace = list(pm = 999999)) %>% 
  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
  # 291,490 obs less than 1; 56,684 obs less than 0 
  mutate(pm = if_else(pm < 1, 0.5, pm))

# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
```

# New variables to use for analysis
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new.rds")
cleaning_summary <- read_xlsx("Input/cleaning_summary.xlsx") %>% 
  select(area, home_winter_id, final_cleaning_pm, sensitivity_analysis_pm)

kids_pm_new <- kids_pm %>% 
  mutate(home_winter_id = as.numeric(home_winter_id)) %>% 
  left_join(cleaning_summary, by = c("area", "home_winter_id")) %>% 
  select(-samplingdate, -samplingtime, -pickupdate, -pickuptime,
         -dtmin, -dtmax, -dtaverage, -qaqccomments, -data_comments,
         -pm_sampling_day, -pm_mean_daily) %>% 
  # create sampling days based on datetime differences
  group_by(area, home_winter_id) %>% 
  mutate(pm_datetime_new = if_else(is.na(pm_datetime_new), 
                                   pm_datetime, pm_datetime_new),
         first_datetime_new = first(pm_datetime_new),
         first_datetime_new = ymd_hms(first_datetime_new)) %>% 
  mutate(datetime_diff = pm_datetime_new - first_datetime_new,
         sampling_day = if_else(datetime_diff <= 86400, 1, 0),
         sampling_day = if_else(datetime_diff > 86400 & datetime_diff <= 172800, 
                                2, sampling_day),
         sampling_day = if_else(datetime_diff > 172800 & datetime_diff <= 259200, 
                                3, sampling_day),
         sampling_day = if_else(datetime_diff > 259200 & datetime_diff <= 345600, 
                                4, sampling_day),
         sampling_day = if_else(datetime_diff > 345600 & datetime_diff <= 432000, 
                                5, sampling_day),
         sampling_day = if_else(datetime_diff > 432000 & datetime_diff <= 518400, 
                                6, sampling_day),
         sampling_day = if_else(datetime_diff > 518400 & datetime_diff <= 604800, 
                                7, sampling_day),
         sampling_day = if_else(datetime_diff > 604800 & datetime_diff <= 691200, 
                                8, sampling_day),
         sampling_day = if_else(datetime_diff > 691200 & datetime_diff <= 777600, 
                                9, sampling_day),
         sampling_day = if_else(datetime_diff > 777600 & datetime_diff <= 864000, 
                                10, sampling_day),
         sampling_day = if_else(datetime_diff > 864000, 
                                11, sampling_day)) %>% 
  # new vars for daily means and % expected observations for each day
  group_by(area, home_winter_id, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm),
         percent_daily_obs_pm = n()/1440) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, pm_datetime_new) %>% 
  select(-first_datetime, -time_diff, -do_not_use)

test <- kids_pm_new %>% 
  filter(percent_daily_obs_pm >= 0.9) %>% 
  summarise(n()/2241695)

# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
# write_csv(kids_pm_new, "Output/kids_pm_new.csv")
```