### 2020-01-13 indoor_temp_rh_cleaning.Rmd

---
title: "KidsAIR indoor temp and humidity cleaning"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 13 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save indoor temp/rh data
```{r}
##### WMT #####
# Load data
temp_rh_initial_wmt <- read_csv("Input/WMT/wmt_indoor_temp_rh_20200113.csv")

wmt_temp_rh <- temp_rh_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         rh_temp_comments = as.character(`REPLACE(REPLACE(Comments`),
         area = "WMT") %>% 
  # replace -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min, rh_temp_comments) %>% 
  arrange(home_winter_id)

# Save as RDS 
# write_rds(wmt_temp_rh, "Output/wmt_temp_rh.rds")

##### NN #####
# Load data
temp_rh_initial_nn <- read_xlsx("Input/NN/nn_indoor_temp_rh_20200113.xlsx")

nn_temp_rh <- temp_rh_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         rh_temp_comments = as.character(IgnoreReason),
         area = "NN") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min, rh_temp_comments) %>% 
  arrange(home_winter_id)

# Save as RDS 
# write_rds(nn_temp_rh, "Output/nn_temp_rh.rds")

##### AK #####
# Load data
temp_rh_initial_ak <- read_xlsx("Input/AK/ak_indoor_temp_rh_20200113.xlsx")

ak_temp_rh <- temp_rh_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         rh_temp_comments = as.character(IgnoreReason),
         area = "AK") %>% 
  # replace -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min, rh_temp_comments) %>% 
  arrange(home_winter_id)

# Save as RDS 
# write_rds(ak_temp_rh, "Output/ak_temp_rh.rds")
```

# Join temp/rh data from different areas
```{r}
wmt_temp_rh <- read_rds("Output/wmt_temp_rh.rds")
nn_temp_rh <- read_rds("Output/nn_temp_rh.rds")
ak_temp_rh <- read_rds("Output/ak_temp_rh.rds")

kids_indoor_temp_rh <- rbind(wmt_temp_rh, nn_temp_rh, ak_temp_rh)

# write_rds(kids_indoor_temp_rh, "Output/kids_indoor_temp_rh.rds")
```

# Clean up all temp/rh data
```{r}
kids_indoor_temp_rh <- read_rds("Output/kids_indoor_temp_rh.rds")

kids_indoor_temp_rh_clean <- kids_indoor_temp_rh

# write_rds(kids_indoor_temp_rh_clean, "Output/kids_indoor_temp_rh_clean.rds")
```