### 2020-01-14 join_format_kids_data.Rmd
cleaning joining datasets begin summarizing data

---
title: "KidsAIR: Join data and format full dataset"
author: "Ethan Walker"
date: "Started 14 Jan 2020, Updated 14 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```
# Use this Markdown to join datasets for analysis

# Load individual datasets
```{r}
# These files are millions of observations long and take time to load/edit
# Make a separate way to join these data and only use when necessary
# Also consider how to join/analyze these data since they were collected
# at different frequencies (20 or 25 min for SUMs; 1 min for PM)
pm_clean <- read_rds("Output/pm_clean.rds")
sums_clean <- read_rds("Output/sums_clean.rds")

# These other files have fewer observations and are easier to work with
# Make a full dataset by joining these datasets together
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
moisture_clean <- read_rds("Output/moisture_clean.rds")
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
home_char_clean <- read_rds("Output/home_char_clean.rds")
home_act_clean <- read_rds("Output/home_act_clean.rds")
```

# Join small datasets
```{r}

```

# Format small dataset
```{r}

```

# Join large datasets
```{r}

```