### 2020-01-16 sums_summary.Rmd

---
title: "Kids SUMs Summary"
author: "Ethan Walker"
date: "Started 9 Jan 2020, Updated 15 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(knitr)
library(lubridate)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
sums_clean <- read_rds("Output/sums_clean.rds")
```

# Summary stats by area (winter)
```{r}
sums_summary <- sums_clean %>% 
  # group by: area, treatment, dtid, day_of_week, Sampling_day
  group_by(area, winter_id) %>% 
  summarize("n" = n(),
            "Mean Temp" = mean(temp_c_sums, na.rm = TRUE), 
            "SD Temp" = sd(temp_c_sums, na.rm = TRUE),
            "Min Temp" = min(temp_c_sums, na.rm = TRUE), 
            "Median Temp" = median(temp_c_sums, na.rm = TRUE),
            "Max Temp" = max(temp_c_sums, na.rm = TRUE)) %>% 
  select(-area, -winter_id) %>% 
  t()

kable(sums_summary, col.names = c("Alaska (1)", "Alaska (2)",
                                  "Navajo (1)", "Navajo (2)",
                                  "Montana (1)", "Montana (2)"), 
      digits = 1, align = "c")
```

# Boxplots for SUMs temp (C) by study area - Winter 1
```{r, fig.height=6, fig.width=5}
boxplots_kids_sums <- sums_clean %>%
  filter(winter_id == 1) %>% 
  ggplot(aes(area, temp_c_sums)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(-50, 100) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = "Temperature (degrees C)") +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_sums
```

# Boxplots for SUMs temp (C) by study area - Winter 2
```{r, fig.height=6, fig.width=5}
boxplots_kids_sums <- sums_clean %>%
  filter(winter_id == 2) %>% 
  ggplot(aes(area, temp_c_sums)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(-50, 100) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = "Temperature (degrees C)") +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_sums
```