### 2020-01-16 indoor_temp_rh_summary.Rmd
minor edit

---
title: "Kids indoor temp/RH summary"
author: "Ethan Walker"
date: "Started 15 Jan 2020, Updated 16 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
```

```{r}
# Load data
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
```

# Summary stats
```{r}
temp_rh_summary <- indoor_temp_rh_clean %>%
  group_by(area) %>% 
  summarize_if(is.numeric, list(mean = mean, sd = sd, min = min, 
                                med = median, max = max), na.rm = TRUE) %>% 
  select(-area) %>% 
  t() 
kable(temp_rh_summary, col.names = c("Alaska", "Navajo", "Montana"), 
      digits = 1, align = "c")
```

```{r, fig.height=6, fig.width=5}
# Boxplot function
boxplots_temp_rh_function <- function(data, variable){

boxplots_temp_rh <- data %>% 
  # change input to plot different temp/RH variable
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  na.exclude() %>% 
  ggplot(aes(area, plot_var)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    #ylim(0, 100) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    #labs(y = "Max indoor RH (%)") +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_temp_rh

}
```

## Current indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_current")
```

## Max indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_max")
```

## Min indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_min")
```

## Current indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_current")
```

## Max indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_max")
```

## Min indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_min")
```