### 2020-01-23 filter_treatment_obs_check.Rmd
data cleaning

---
title: "KidsAIR: treatment filtering and missing data check"
author: "Ethan Walker"
date: "Started 21 Jan 2020, Updated 23 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load individual datasets
```{r}
pm_clean <- read_rds("Output/pm_clean.rds")
sums_clean <- read_rds("Output/sums_clean.rds")

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
moisture_clean <- read_rds("Output/moisture_clean.rds")
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
home_char_clean <- read_rds("Output/home_char_clean.rds")
home_act_clean <- read_rds("Output/home_act_clean.rds")
```

# Filter winter 1 and specific treatment homes from PM dataset
## These homes will be the focus of the initial exposure abstract/paper
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

# 88 Winter 1 education homes have PM data
pm_homes_education <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Education") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

# 86 Winter 1 filter homes have PM data
pm_homes_filter <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Filter") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)
```

#### Join selected homes from above with other datasets and filter homes of interest ####

# PM data
## 93 placebo homes from Winter 1 have PM data
## 88 education homes from Winter 1 have PM data
## 86 filter homes from Winter 1 have PM data
## All data is accounted for; missing data is truly missing from data collection issues
```{r}
pm_filtered <- pm_clean %>% 
  left_join(pm_homes_placebo, by = c("area", "home")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  ungroup() %>% 
  arrange(area, home)

pm_obs <- pm_clean %>% 
  left_join(pm_homes_placebo, by = c("area", "home")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  group_by(area, home) %>% 
  summarize(n())
```

# SUMs data
## 80/93 placebo homes from Winter 1 have SUMs data
## 77/88 education homes from Winter 1 have SUMs data
## 70/86 filter homes from Winter 1 have SUMs data
## All data is accounted for; missing data is truly missing from data collection issues
```{r}
sums_check <- sums_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  group_by(area, home) %>% 
  summarize(n())

sums_filtered <- sums_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Demographics data
## 93/93 placebo homes from Winter 1 have demographic data
## 88/88 education homes from Winter 1 have demographic data
## 86/86 filter homes from Winter 1 have demographic data
#### Some missing data within individual homes/rows; will be assessed in further data cleaning ####
#### Check for different answers across Winter 1 and Winter 2 ####
```{r}
demographics_check1 <- demographics_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  group_by(area, home) %>% 
  summarize(n())

demographics_check2 <- demographics_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  group_by(area, home) %>% 
  summarize(n()) %>% 
  mutate(dem_filtered = "true") %>% 
  # change to correct joining data of interest
  full_join(pm_homes_placebo, by = c("area", "home"))

demographics_filtered <- demographics_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Moisture data
## 91/93 placebo homes from Winter 1 have moisture data
## 83/88 education homes from Winter 1 have moisture data
## 83/86 filter homes from Winter 1 have moisture data
```{r}
moisture_check <- moisture_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(moisture_ave)) %>% 
  group_by(area, home) %>% 
  summarize(n())

moisture_filtered <- moisture_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(moisture_ave)) %>% 
  ungroup() %>% 
  arrange(area, home)

# Quick data check for odd values
min(moisture_filtered$moisture_ave)
max(moisture_filtered$moisture_ave)
```

# Indoor temp/rh data
## 75/93 placebo homes from Winter 1 have temp/rh data
## 70/88 education homes from Winter 1 have temp/rh data
## 76/86 filter homes from Winter 1 have temp/rh data
#### Some further missing data within individual homes/rows ####
```{r}
indoor_temp_rh_check <- indoor_temp_rh_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  mutate(na_filter = if_else(is.na(rh_indoor_current) & is.na(rh_indoor_max) & is.na(rh_indoor_min) &
                             is.na(temp_indoor_current) & is.na(temp_indoor_max) & is.na(temp_indoor_min),
                             1, 0)) %>% 
  filter(na_filter == 0) %>% 
  group_by(area, home) %>% 
  summarize(n())

indoor_temp_rh_filtered <- indoor_temp_rh_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  mutate(na_filter = if_else(is.na(rh_indoor_current) & is.na(rh_indoor_max) & is.na(rh_indoor_min) &
                             is.na(temp_indoor_current) & is.na(temp_indoor_max) & is.na(temp_indoor_min),
                             1, 0)) %>% 
  filter(na_filter == 0) %>% 
  ungroup() %>% 
  arrange(area, home)

# Quick data check for odd values
check_vars <- list(indoor_temp_rh_filtered$rh_indoor_current, indoor_temp_rh_filtered$rh_indoor_max, 
                   indoor_temp_rh_filtered$rh_indoor_min, indoor_temp_rh_filtered$temp_indoor_current, 
                   indoor_temp_rh_filtered$temp_indoor_max, indoor_temp_rh_filtered$temp_indoor_min)

lapply(check_vars, min, na.rm = TRUE)
lapply(check_vars, max, na.rm = TRUE)
```

# Stove use data
## 92/93 placebo homes from Winter 1 have stove use data
## 86/88 education homes from Winter 1 have stove use data
## 85/86 filter homes from Winter 1 have stove use data
#### Some further missing data within individual homes/rows ####
#### 5 homes reported using 20+ cords per year; confirmed by checking paper logs ####
```{r}
stove_use_check <- stove_use_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  mutate(na_filter = if_else(is.na(burn_level) & is.na(wood_type) & 
                             is.na(wood_collect) & is.na(wood_collect_method),
                             1, 0)) %>% 
  filter(na_filter == 0) %>%
  group_by(area, home) %>% 
  summarize(n())

stove_use_filtered <- stove_use_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  mutate(na_filter = if_else(is.na(burn_level) & is.na(wood_type) & 
                             is.na(wood_collect) & is.na(wood_collect_method),
                             1, 0)) %>% 
  filter(na_filter == 0) %>%
  ungroup() %>% 
  arrange(area, home)

# Quick data check for odd values
min(stove_use_filtered$wood_cords, na.rm = TRUE)
max(stove_use_filtered$wood_cords, na.rm = TRUE)
```

# Stove grades data
## 90/93 placebo homes have stove grades data
## 83/88 education homes have stove grades data
## 79/86 filter homes have stove grades data
#### If Winter 1 was missing, the grade for Winter 2 was used ####
```{r}
stove_grades_check <- stove_grades_clean %>% 
  # change to correct joining data of interest
 left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  filter(!is.na(stove_grade)) %>% 
  group_by(area, home) %>% 
  summarize(n())

stove_grades_filtered <- stove_grades_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  filter(!is.na(stove_grade)) %>% 
  ungroup() %>% 
  arrange(area, home)

table(stove_grades_filtered$stove_grade)
```

# Home characteristics data
## Using answers from Winter 2 if Winter 1 = NA
## 92/93 placebo homes have home characteristics data
## 86/88 education homes have home characteristics data
## 85/86 filter homes have home characteristics data
#### Need to assess differences between Winter 1 and Winter 2 ####
```{r}
home_char_check <- home_char_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  group_by(area, home) %>% 
  mutate_all(funs(if_else(is.na(.), lead(.), .))) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  mutate(na_filter = if_else(is.na(residents_smoke) & is.na(home_type) & 
                             is.na(primary_heat) & is.na(stove_age),
                             1, 0)) %>%
  filter(na_filter == 0) %>%
  group_by(area, home) %>% 
  summarize(n())

home_char_filtered <- home_char_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  arrange(area, home, winter_id) %>% 
  group_by(area, home) %>% 
  mutate_all(funs(if_else(is.na(.), lead(.), .))) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  mutate(na_filter = if_else(is.na(residents_smoke) & is.na(home_type) & 
                             is.na(primary_heat) & is.na(stove_age),
                             1, 0)) %>%
  filter(na_filter == 0) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Home activity data
## 93/93 placebo homes from Winter 1 have home activity data
### 92/93 have 6 observations; 1 has 5 observations
## 88/88 education homes from Winter 1 have home activity data
### 84/88 have 6 observations; 4 have 5 observations
## 86/86 filter homes from Winter 1 have home activity data
### 85/86 have 6 observations; 1 has 5 observations
```{r}
home_act_check <- home_act_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  group_by(area, home) %>% 
  summarize(n())

home_act_filtered <- home_act_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  ungroup() %>% 
  arrange(area, home, home_act_day)
```