### 2020-01-27 demographics_summary.Rmd
Summary tables for abstacts/papers

---
title: "KidsAIR: household demographics summary"
author: "Ethan Walker"
date: "Started 27 Jan 2020, Updated 27 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load individual datasets
```{r}
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
pm_clean <- read_rds("Output/pm_clean.rds")
```

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

demographics_filtered <- demographics_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Summarize demographics data
```{r}
# Summarize numeric vars
kids_dems <- demographics_filtered %>%  
  mutate(total_residents = as.numeric(total_residents),
         residents_under_five = as.numeric(residents_under_five)) %>% 
  distinct(home, .keep_all = TRUE) %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(total_residents, na.rm = TRUE), 
            "SD" = sd(total_residents, na.rm = TRUE),
            "min" = min(total_residents, na.rm = TRUE), 
            "max" = max(total_residents, na.rm = TRUE))
kable(kids_dems, digits = 1, align = "c")

# Summarize factor vars
alaska_dems <- demographics_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "AK") %>% 
  group_by(age_child_nov) %>% 
  #summarize(n = n(), percent = (n()/10)*100)
  summarize(n = n(), percent = (n()/18)*100)
alaska_dems

navajo_dems <- demographics_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "NN") %>% 
  group_by(age_child_nov) %>% 
  #summarize(n = n(), percent = (n()/23)*100)
  summarize(n = n(), percent = (n()/32)*100)
navajo_dems

montana_dems <- demographics_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "WMT") %>% 
  group_by(age_child_nov) %>% 
  #summarize(n = n(), percent = (n()/60)*100)
  summarize(n = n(), percent = (n()/87)*100)
montana_dems

## Select specific var to assess individually/group by area
table(demographics_filtered$education, demographics_filtered$area, useNA = "ifany")
```