### 2020-03-24 exposure_initial_analysis.Rmd
merge datasets

---
title: "KidsAIR: initial exposure analysis"
author: "Ethan Walker"
date: "Started 2 Feb 2020, Updated 20 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(lme4)
library(lmerTest)
```

# Load individual datasets
```{r}
pm_clean <- read_rds("Output/pm_clean.rds")
sums_clean <- read_rds("Output/sums_clean.rds")

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
moisture_clean <- read_rds("Output/moisture_clean.rds")
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
home_char_clean <- read_rds("Output/home_char_clean.rds")
home_act_clean <- read_rds("Output/home_act_clean.rds")

# Mean PM for each placebo home over winter 1
pm_home_mean <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(area, home) %>% 
  summarize(mean_pm = mean(pm, na.rm = TRUE))

exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")

exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds") 
```

# Demographics
```{r}
filter_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean))

# Variables to use:
## education_3level, income_3level, race_parent_2level, race_child_2level
## total_residents_2level, residents_under5_2level, resident_smokes

table(filter_data$residents_smoke)
median(filter_data$home_furry, na.rm = TRUE)

pm_demographics_results <- lm(log(pm_mean) ~ residents_smoke, data = filter_data)
summary(pm_demographics_results)
plot(pm_demographics_results)
```

# Home characteristics
```{r}
filter_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean))

# Variables to use:
## home_type_2level, home_year_built_2level, home_sqft_2level, home_floors_2level
## home_bedrooms_2level, home_windows_2level, home_pets_2level, home_furry_2level
## secondary_heat_3level, stove_age_3level, stove_cert, chimney_clean_3level,
## home_damp, home_mold, home_fireplace, home_furnace, home_insert

table(filter_data$home_insert)
median(filter_data$home_furry, na.rm = TRUE)

pm_home_char_results <- lm(log(pm_mean) ~ home_insert, data = filter_data)
summary(pm_home_char_results)
plot(pm_home_char_results)
```

# Stove use
```{r}
filter_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean))

# Variables to use:
## burn_level_3level, wood_collect_2level, wood_collect_method_2level,
## wood_cords

table(filter_data$home_insert)
median(filter_data$wood_cords, na.rm = TRUE)

pm_stove_use_results <- lm(log(pm_mean) ~ wood_collect_method_2level, data = filter_data)
summary(pm_stove_use_results)
plot(pm_stove_use_results)
pm_stove_use_plot <- pm_stove_use %>% 
  ggplot() + 
    geom_boxplot(aes(burn_level, mean_pm))
pm_stove_use_plot
```

# Stove grades
```{r}
filter_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean))

# Variables to use:
## stove_grade_3level

table(filter_data$stove_grade_3level)

pm_stove_grades_results <- lm(log(pm_mean) ~ stove_grade_3level, data = filter_data)
summary(pm_stove_grades_results)
anova(pm_stove_grades_results)
plot(pm_stove_grades_results)
pm_stove_grades_plot <- pm_stove_grades %>% 
  ggplot() + 
    geom_boxplot(aes(stove_grade, mean_pm))
pm_stove_grades_plot
```

# Home activity
```{r}
filter_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean_sampling_day))

# Variables to use:
## home_act_gas, home_act_elect, home_act_propane, home_act_oil, home_act_other,
## home_act_smoke, home_act_incense, home_act_candle, home_act_windows,
## home_act_door, home_act_sweep

table(filter_data$home_act_sweep)

pm_home_act_results <- lmer(log(pm_mean_sampling_day) ~ home_act_sweep + 
                              (1 | home), data = filter_data)
summary(pm_home_act_results)
plot(pm_home_act_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_home_act_results), main = "QQ Plot")
```

# Wood moisture content
```{r}
filter_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(!is.na(moisture_ave)) %>% 
  select(sample_date, pm_first_datetime, area, home, 
         moisture_ave, pm_mean_sampling_period) %>% 
  separate(pm_first_datetime, c("pm_first_date", "first_time"), sep = " ") %>% 
  group_by(area, home) %>% 
  mutate(moisture_date = ymd(sample_date),
         first_date = ymd(pm_first_date),
         time_diff = moisture_date - first_date,
         time_diff = as.numeric(abs(time_diff))) %>% 
  group_by(home) %>% 
  mutate(time_diff_rank = if_else(time_diff == min(time_diff, na.rm = TRUE), 1, 0)) %>%
  ungroup() %>% 
  filter(time_diff_rank == 1) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean_sampling_period))

# Variables to use:
## moisture_ave, mean_moisture_5

table(filter_data$moisture_cat)
median(filter_data$moisture_ave, na.rm = TRUE)

pm_moisture_results <- lm(log(pm_mean_sampling_period) ~ 
                            moisture_ave, data = filter_data)
summary(pm_moisture_results)
plot(pm_moisture_results)
```

# Indoor temp/Rh
```{r}
pm_temp_rh <- pm_home_mean %>% 
  left_join(indoor_temp_rh_clean, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(temp_max_5 = temp_indoor_max/5,
         temp_min_5 = temp_indoor_min/5,
         rh_max_5 = rh_indoor_max/5,
         rh_min_5 = rh_indoor_min/5,
         temp_max_cat = cut(temp_indoor_max, breaks = c(0, 30, 45),
                                labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_cat = cut(temp_indoor_min, breaks = c(0, 16.5, 50),
                                labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_cat = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_cat = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                labels = c("<20", "20+"),
                                 right = FALSE))

table(pm_temp_rh$rh_min_cat)
median(pm_temp_rh$rh_indoor_min, na.rm = TRUE)
range(pm_temp_rh$rh_indoor_max, na.rm = TRUE)

pm_home_act_results <- lm(log(mean_pm) ~ rh_min_cat, data = filter_data)
summary(pm_home_act_results)
plot(pm_home_act_results)
```

# SUMs - experimenting with several methods of looking at associations
```{r}
#### Join PM dataset full SUMs dataset by SUMs sampling times

pm_join <- pm_clean %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first_datetime_new,
         sample_datetime = round_date(sample_datetime, unit = "minute"),
         pm_rolling_20 = rollmean(pm, 20, fill = "extend")) %>% 
  ungroup() %>% 
  arrange(area, home, sample_datetime)

sums_join <- sums_clean %>% 
  mutate(sample_datetime = ymd_hms(datetime_sums)) %>% 
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  mutate(sample_datetime = round_date(sample_datetime, unit = "minute")) %>% 
  ungroup() %>% 
  arrange(area, home, sample_datetime)

pm_sums_joined <- sums_join %>% 
  left_join(pm_join, by = c("area", "home", "sample_datetime")) %>% 
  filter(pm_rolling_20 < 1000) %>% 
  filter(!is.na(temp_c_sums)) %>% 
  filter(!is.na(pm_rolling_20)) %>% 
  mutate(temp_c_sums_5 = temp_c_sums/5)

pm_sums_count <- pm_sums_joined %>% 
  count(home)

cor.test(pm_sums_joined$temp_c_sums, pm_sums_joined$pm_rolling_20, method = "spearman")
pm_sums_model <- lmer(log(pm_rolling_20) ~ temp_c_sums_5 + (1 | home), data = pm_sums_joined)

summary(pm_sums_model)
exp((0.146291)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

pm_sums_plot <- pm_sums_joined %>% 
  ggplot() + 
    geom_point(aes(log(pm_rolling_20), temp_c_sums)) +
    geom_smooth(aes(log(pm_rolling_20), temp_c_sums)) +
    theme_minimal()
pm_sums_plot

#### Join PM dataset with SUMs dataset by sampling day

pm_join <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  mutate(pm = if_else(pm > 500, 500, pm),
         pm_mean = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm, na.rm = TRUE)) %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  group_by(area, home, sampling_day) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first(sample_datetime)) %>% 
  ungroup() %>% 
  arrange(area, home, sample_datetime) %>% 
  select(area, home, home_winter_id, dtid, day_of_week, pm_total_observations,
         pm_mean, sampling_day, pm_mean_daily, pm_first_datetime, pm_last_datetime) %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sampling_day)

sums_join <- sums_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  select(area, home, winter_id, ibid, treatment, datetime_sums, temp_c_sums,)

pm_sums_joined <- sums_join %>% 
  left_join(pm_join, by = c("area", "home")) %>% 
  arrange(area, home, datetime_sums) %>% 
  mutate(filter_var = if_else(datetime_sums >= pm_first_datetime &
                              datetime_sums <= pm_last_datetime, 1, 0)) %>% 
  filter(filter_var == 1) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(temp_perc_25 = if_else(temp_c_sums >= 30, 1, 0)) %>% 
  arrange(area, home, sampling_day) %>% 
  mutate(sums_mean_daily = mean(temp_c_sums, na.rm = TRUE),
         sums_mean_daily_5 = sums_mean_daily/5) %>% 
  group_by(area, home, home_winter_id, sampling_day, pm_mean, 
           pm_mean_daily, sums_mean_daily) %>% 
  summarize(temp_perc_25 = (sum(temp_perc_25)/n()*100)) %>% 
  mutate(temp_perc_25_cat = cut(temp_perc_25, breaks = c(0, 10, 200),
                                labels = c("< 50%", "50% +"),
                                 right = FALSE)) %>%
  distinct(pm_mean_daily, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sampling_day)

pm_sums_count <- pm_sums_joined %>% 
  count(home)
median(pm_sums_joined$temp_perc_25)
table(pm_sums_joined$temp_perc_25_cat)

cor.test(pm_sums_joined$sums_mean_daily, pm_sums_joined$pm_mean_daily, method = "spearman")
pm_sums_model <- lmer(pm_mean_daily ~ temp_perc_25_cat + (1 | home), data = pm_sums_joined)
pm_sums_model <- lmer(log(pm_mean_daily) ~ temp_perc_25_cat + (1 | home), data = pm_sums_joined)

summary(pm_sums_model)
exp((0.146291)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

#### Join PM dataset with SUMs events datasets

# Mean PM for each placebo home over winter 1
pm_home_mean_date <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  separate(pm_datetime_new, c("date_sums", "pm_time"), sep = " ", remove = TRUE) %>% 
  group_by(area, home, date_sums) %>% 
  summarize(mean_pm = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, date_sums) %>% 
  distinct(date_sums, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, date_sums)

sums_temp_change <- sums_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home, date_sums) %>% 
  mutate(time_diff = (lead(datetime_sums) - datetime_sums)/60) %>% 
  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
  replace_with_na(replace = list(lead_temp = 999)) %>% 
  mutate(temp_diff = lead_temp - temp_c_sums,
         temp_diff_abs = abs(lead_temp - temp_c_sums),
         # change temp in the following line
         temp_diff_check = if_else(temp_diff >= 8, 1, 0),
         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
                              temp_diff_check),
         heat_event = if_else(is.na(temp_diff_check), 1,
                              heat_event)) %>% 
  ungroup() %>% 
  mutate(date_sums = as.character(date_sums)) %>% 
  group_by(area, home, date_sums) %>% 
  summarize(sums_events = sum(heat_event, na.rm = TRUE)) %>% 
  ungroup() %>% 
  arrange(area, home, date_sums)

pm_sums_events_joined <- pm_home_mean_date %>% 
  left_join(sums_temp_change, by = c("area", "home", "date_sums")) %>% 
  arrange(area, home, date_sums) %>% 
  #mutate(sums_events = factor(sums_events))
  mutate(sums_events_cat = cut(sums_events, breaks = c(0, 2, 20),
                           labels = c("<2", "2+"),
                                 right = FALSE))

median(pm_sums_events_joined$sums_events, na.rm = TRUE)
table(pm_sums_events_joined$sums_events_cat)

pm_sums_model <- lmer(mean_pm ~ sums_events_cat + (1 | home), data = filter_data)
pm_sums_model <- lmer(log(mean_pm) ~ sums_events_cat + (1 | home), data = filter_data)

summary(pm_sums_model)
#exp((0.146291)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

pm_sums_plot <- pm_sums_events_joined %>% 
  filter(!is.na(sums_events)) %>% 
  ggplot() + 
    geom_boxplot(aes(sums_events_cat, log(mean_pm))) +
    theme_minimal()
pm_sums_plot

```