### 2020-03-30 ambient_initial.Rmd
ambient data updates

---
title: "KidsAIR: ambient data"
author: "Ethan Walker"
date: "Started 30 March 2020, Updated 30 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(ggmap)
library(maps)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data

wmt_zipcodes <- read_csv("Input/WMT/wmt_zipcodes.csv") %>% 
  rename_all(tolower)

nn_zipcodes <- read_csv("Input/NN/nn_zipcodes.csv") %>% 
  rename_all(tolower)

ak_zipcodes <- read_csv("Input/AK/ak_zipcodes.csv") %>% 
  rename_all(tolower)

us_zipcodes <- read_tsv("Input/us_zipcodes/us_zipcodes.txt", 
                        col_names = c("country", "zip", "city", "state",
                                      "state_abbr", "trash1", "trash2", "trash3", 
                                      "trash4", "lat", "long", "trash5")) %>% 
  select(-trash1:-trash4, -trash5)

wmt_map <- map_data("state") %>% 
  filter(region == "montana" | region == "idaho")

nn_map <- map_data("state") %>% 
  filter(region == "arizona" | region == "utah" |
         region == "new mexico" | region == "colorado")

ak_map <- map_data("state") %>% 
  filter(region == "alaska")

montana_temps <- read_csv("Input/ambient/montana_temp.txt") %>% 
  rename_all(tolower) %>% 
  mutate(station = `stn---`,
         temp_date = yearmoda,
         mean_temp = temp,
         mean_dewp = dewp,
         mean_pressure = stp,
         mean_visib_miles = visib,
         mean_wind_knots = wdsp,
         max_wind_knots = mxspd,
         max_temp = max,
         min_temp = min,
         precip = prcp,
         snow_depth = sndp) %>% 
  select(station:snow_depth)

us_temp_stations <- read_table("Input/ambient/temp_station_list.txt", skip = 20,
                            col_names = c("station", "wban", "station_names",
                                          "country", "state", "call", "lat",
                                          "long", "elevation", "begin_date",
                                          "end_date")) %>% 
  filter(country == "US") %>% 
  mutate(lat = as.numeric(lat),
         long = as.numeric(long),
         elevation = as.numeric(elevation),
         begin_date = ymd(begin_date),
         end_date = ymd(end_date))

pm_stations <- read_xlsx("Input/ambient/pm_station_list.xlsx")
```

# Mapping zipcodes, weather stations, pm stations - WMT
```{r}
wmt_zipcodes_new <- wmt_zipcodes %>% 
  left_join(us_zipcodes, by = "zip") %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  mutate(lat_zip = lat,
         long_zip = long)

#Plot temp stations
wmt_temp_stations <- us_temp_stations %>% 
  filter(state == "MT" | state == "ID") %>% 
  mutate(lat_temp = lat,
         long_temp = long) %>% 
  distinct(station, .keep_all = TRUE) %>% 
  full_join(wmt_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_temp),
         long_diff = abs(long_zip - long_temp),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(home))

wmt_plot <- wmt_zipcodes_new %>% 
  ggplot() +
    geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
               colour="black", fill = "white") +
    geom_point(data=wmt_temp_stations, 
             aes(long_zip, lat_zip, color = station_names), size = 3) +
    geom_point(data=wmt_temp_stations, 
             aes(long_temp, lat_temp, color = station_names), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette)
wmt_plot

#Plot PM stations
wmt_pm_stations <- pm_stations %>% 
  filter(area == "wmt") %>% 
  full_join(wmt_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_pm),
         long_diff = abs(long_zip - long_pm),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(home))

wmt_plot <- wmt_zipcodes_new %>% 
  ggplot() +
    geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
               colour="black", fill = "white") +
    geom_point(data=wmt_pm_stations, 
             aes(long_zip, lat_zip, color = site), size = 3) +
    geom_point(data=wmt_pm_stations, 
             aes(long_pm, lat_pm, color = site), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette)
wmt_plot
```

# Mapping zipcodes, weather stations, pm stations - NN
```{r}
nn_zipcodes_new <- nn_zipcodes %>% 
  left_join(us_zipcodes, by = "zip") %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  mutate(lat_zip = lat,
         long_zip = long)

#Plot temp stations
nn_temp_stations <- us_temp_stations %>% 
  filter(state == "AZ" | state == "NM" |
         state == "UT" | state == "CO") %>% 
  mutate(lat_temp = lat,
         long_temp = long) %>% 
  distinct(station, .keep_all = TRUE) %>% 
  full_join(nn_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_temp),
         long_diff = abs(long_zip - long_temp),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(home))

nn_plot_temp_stations <- wmt_zipcodes_new %>% 
  ggplot() +
    geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
               colour="black", fill = "white") +
    geom_point(data=nn_temp_stations, 
             aes(long_zip, lat_zip, color = station_names), size = 3) +
    geom_point(data=nn_temp_stations, 
             aes(long_temp, lat_temp, color = station_names), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette)
nn_plot_temp_stations

#Plot PM stations
nn_pm_stations <- pm_stations %>% 
  filter(area == "nn") %>% 
  full_join(nn_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_pm),
         long_diff = abs(long_zip - long_pm),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(home))

nn_plot <- nn_zipcodes_new %>% 
  ggplot() +
    geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
               colour="black", fill = "white") +
    geom_point(data=nn_pm_stations, 
             aes(long_zip, lat_zip, color = site), size = 3) +
    geom_point(data=nn_pm_stations, 
             aes(long_pm, lat_pm, color = site), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette)
nn_plot
```

# Mapping zipcodes, weather stations, pm stations - WMT
```{r}
ak_zipcodes_new <- ak_zipcodes %>% 
  left_join(us_zipcodes, by = "zip") %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  mutate(lat_zip = lat,
         long_zip = long)

#Plot temp stations
ak_temp_stations <- us_temp_stations %>% 
  filter(state == "AK") %>% 
  mutate(lat_temp = lat,
         long_temp = long) %>% 
  distinct(station, .keep_all = TRUE) %>% 
  full_join(ak_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_temp),
         long_diff = abs(long_zip - long_temp),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE)

zipcodes_map <- wmt_zipcodes_new %>% 
  ggplot() +
    geom_point(data=ak_temp_stations, 
             aes(long_zip, lat_zip, color = station_names), size = 3) +
    geom_point(data=ak_temp_stations, 
             aes(long_temp, lat_temp, color = station_names), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette) 
zipcodes_map

#Plot PM stations
ak_pm_stations <- pm_stations %>% 
  filter(area == "ak") %>% 
  full_join(ak_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_pm),
         long_diff = abs(long_zip - long_pm),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(home))

ak_plot <- ak_zipcodes_new %>% 
  ggplot() +
    geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
               colour="black", fill = "white") +
    geom_point(data=ak_pm_stations, 
             aes(long_zip, lat_zip, color = site), size = 3) +
    geom_point(data=ak_pm_stations, 
             aes(long_pm, lat_pm, color = site), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette)
ak_plot
```