### 2020-03-30 exposure_initial_analysis.Rmd

---
title: "KidsAIR: initial exposure analysis"
author: "Ethan Walker"
date: "Started 2 Feb 2020, Updated 30 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(ggmap)
```

# Load individual datasets
```{r}
exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long_new.rds")

exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")

exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds")

sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")

health_exposure_data_sampling_day <- read_rds("Output/health_exposure_data_sampling_day.rds")
```

# Demographics
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## education_3level, income_3level, race_parent_2level, race_child_2level
## total_residents_2level, residents_under5_2level, resident_smokes

table(analysis_data$residents_smoke)
median(analysis_data$home_furry, na.rm = TRUE)

pm_demographics_results <- lm(log(pm_mean_sampling_period) ~ residents_smoke, 
                              data = analysis_data)
summary(pm_demographics_results)
tidy(pm_demographics_results)
plot(pm_demographics_results)
```

# Home characteristics
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## home_type_2level, home_year_built_2level, home_sqft_2level, home_floors_2level
## home_bedrooms_2level, home_windows_2level, home_pets_2level, home_furry_2level
## secondary_heat_3level, stove_age_3level, stove_cert, chimney_clean_3level,
## home_damp, home_mold, home_fireplace, home_furnace, home_insert

table(analysis_data$home_insert)
median(analysis_data$home_furry, na.rm = TRUE)

pm_home_char_results <- lm(log(pm_mean_sampling_period) ~ home_insert, 
                           data = analysis_data)
summary(pm_home_char_results)
tidy(pm_home_char_results)
plot(pm_home_char_results)
```

# Stove use
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## burn_level_3level, wood_collect_2level, wood_collect_method_2level,
## wood_cords

table(analysis_data$burn_level_3level)
median(analysis_data$wood_cords, na.rm = TRUE)

pm_stove_use_results <- lm(log(pm_mean_sampling_period) ~ wood_collect_method_2level, 
                           data = analysis_data)
summary(pm_stove_use_results)
tidy(pm_stove_use_results)
plot(pm_stove_use_results)
pm_stove_use_plot <- pm_stove_use %>% 
  ggplot() + 
    geom_boxplot(aes(burn_level, mean_pm))
pm_stove_use_plot
```

# Stove grades
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## stove_grade_3level

table(analysis_data$stove_grade_3level)

pm_stove_grades_results <- lm(log(pm_mean_sampling_period) ~ stove_grade_3level, 
                              data = analysis_data)
summary(pm_stove_grades_results)
tidy(pm_stove_grades_results)
anova(pm_stove_grades_results)
plot(pm_stove_grades_results)
pm_stove_grades_plot <- pm_stove_grades %>% 
  ggplot() + 
    geom_boxplot(aes(stove_grade, mean_pm))
pm_stove_grades_plot
```

# Home activity
```{r}
analysis_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## home_act_gas, home_act_elect, home_act_propane, home_act_oil, home_act_other,
## home_act_smoke, home_act_incense, home_act_candle, home_act_windows,
## home_act_door, home_act_sweep

table(analysis_data$home_act_sweep)

pm_home_act_results <- lmer(log(pm_mean_sampling_day) ~ home_act_sweep + 
                              (1 | home), data = analysis_data)
summary(pm_home_act_results)
tidy(pm_home_act_results)
plot(pm_home_act_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_home_act_results), main = "QQ Plot")
```

# Wood moisture content
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE))

# Variables to use:
## moisture_closest, moisture_closest_2level
## mean_moisture_winter, moisture_winter_2level

table(analysis_data$moisture_closest_2level)
median(analysis_data$mean_moisture_winter, na.rm = TRUE)

pm_moisture_results <- lm(log(pm_mean_sampling_period) ~ 
                            moisture_winter_2level, data = analysis_data)
summary(pm_moisture_results)
plot(pm_moisture_results)
```

# Indoor temp/RH
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE))

# Variables to use:
## rh_min_5, rh_max_5, temp_min_5, temp_max_5
## temp_max_2level, temp_min_2level,rh_max_2level ,rh_min_2level

table(analysis_data$temp_max_2level)
median(analysis_data$rh_indoor_min, na.rm = TRUE)

pm_temp_rh_results <- lm(log(pm_mean_sampling_period) ~ temp_max_2level, 
                         data = analysis_data)
summary(pm_temp_rh_results)
tidy(pm_temp_rh_results)
plot(pm_temp_rh_results)
```

# SUMs - by sampling day
```{r}
analysis_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  mutate(sums_mean_sampling_day_5 = sums_mean_sampling_day/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  ungroup()

# Variables to use:
## sums_mean_sampling_day_5, temp_perc_25_2level
## temp_perc_27_2level, temp_perc_30_2level

table(analysis_data$temp_perc_30_2level)
median(analysis_data$temp_perc_30, na.rm = TRUE)

pm_sums_results <- lmer(log(pm_mean_sampling_day) ~ temp_perc_30_2level + 
                              (1 | home), data = analysis_data)
pm_sums_results <- lmer(pm_mean_sampling_day ~ temp_perc_30_2level + 
                              (1 | home), data = analysis_data)
summary(pm_sums_results)
tidy(pm_sums_results)
plot(pm_sums_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_results), main = "QQ Plot")
```

# SUMs - by sampling period
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  ungroup()

# Variables to use:
## sums_mean_sampling_period_5, temp_perc_25_2level
## temp_perc_27_2level, temp_perc_30_2level

table(analysis_data$temp_perc_25_2level)
median(analysis_data$temp_perc_30, na.rm = TRUE)

pm_sums_results <- lm(log(pm_mean_sampling_period) ~ sums_mean_sampling_period_5, 
                      data = analysis_data)
pm_sums_results <- lm(pm_mean_sampling_period ~ sums_mean_sampling_period_5, 
                      data = analysis_data)
summary(pm_sums_results)
tidy(pm_sums_results)
plot(pm_sums_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_results), main = "QQ Plot")
```

# SUMs - mean PM per SUMs interval
```{r}
analysis_data <- sums_rolling_pm_data %>% 
  mutate(temp_c_sums_5 = temp_c_sums/5)

cor.test(analysis_data$temp_c_sums, analysis_data$pm_rolling_20, method = "spearman")
pm_sums_model <- lmer(log(pm_rolling_20) ~ temp_c_sums_5 + (1 | home), 
                      data = analysis_data)
pm_sums_model <- lmer(pm_rolling_20 ~ temp_c_sums_5 + (1 | home), 
                      data = analysis_data)

summary(pm_sums_model)
tidy(pm_sums_model)
exp((0.1496794)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

pm_sums_plot <- analysis_data %>% 
  ggplot() + 
    geom_point(aes(log(pm_rolling_20), temp_c_sums)) +
    geom_smooth(aes(log(pm_rolling_20), temp_c_sums)) +
    theme_minimal()
pm_sums_plot
```

# SUMs - events per sampling day 
```{r}
sums_temp_change <- sums_rolling_pm_data %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home, sampling_day) %>% 
  distinct(datetime_sums, .keep_all = TRUE) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(time_diff = (lead(datetime_sums) - datetime_sums)) %>% 
  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
  replace_with_na(replace = list(lead_temp = 999)) %>% 
  mutate(temp_diff = lead_temp - temp_c_sums,
         # change temp in the following line
         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
                              temp_diff_check),
         heat_event = if_else(is.na(temp_diff_check), 1,
                              heat_event)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(sums_events = sum(heat_event, na.rm = TRUE),
         sums_events_cat = cut(sums_events, breaks = c(0, 2, 50),
                           labels = c("<2", "2+"),
                                 right = FALSE)) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup() %>% 
  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
         time_diff:sums_events_cat) %>% 
  arrange(area, home, sample_date)

median(sums_temp_change$sums_events, na.rm = TRUE)
table(sums_temp_change$sums_events_cat)

pm_sums_model <- lmer(pm_mean_sampling_day ~ sums_events_cat + 
                        (1 | home), data = sums_temp_change)
pm_sums_model <- lmer(log(pm_mean_sampling_day) ~ sums_events_cat + 
                        (1 | home), data = sums_temp_change)

summary(pm_sums_model)
tidy(pm_sums_model)
#exp((0.146291)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

pm_sums_plot <- sums_temp_change %>% 
  filter(!is.na(sums_events)) %>% 
  ggplot() + 
    geom_boxplot(aes(sums_events_cat, log(pm_mean_sampling_day))) +
    theme_minimal()
pm_sums_plot
```