### 2020-04-03 ambient_summary.Rmd

---
title: "KidsAIR: ambient temperature and PM analysis"
author: "Ethan Walker"
date: "Started 1 April 2020, Updated 3 April 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE,
                      fig.height = 6, fig.width = 8)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(ggmap)
library(maps)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long_new.rds")

exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")

exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds")

sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")

health_exposure_data_sampling_day <- read_rds("Output/health_exposure_data_sampling_day.rds")
```

```{r, eval=FALSE, include=FALSE}
# data from each sampling day
pm_summary <- exposure_analysis_data_long %>% 
  group_by(area) %>% 
  summarize(mean_amb_pm = mean(amb_pm_24hr, na.rm = TRUE))
pm_summary

temp_summary <- exposure_analysis_data_long %>% 
  group_by(area) %>% 
  summarize(mean_amb_temp = mean(mean_temp, na.rm = TRUE))
temp_summary

# data from indoor PM sampling
pm_summary <- exposure_analysis_data_medium %>% 
  group_by(area) %>% 
  summarize(mean_amb_pm = mean(amb_pm_24hr, na.rm = TRUE))
pm_summary

temp_summary <- exposure_analysis_data_medium %>% 
  group_by(area) %>% 
  summarize(mean_amb_temp = mean(mean_temp, na.rm = TRUE))
temp_summary
```

```{r}
boxplots_amb_pm <- exposure_analysis_data_medium %>% 
  #filter(winter_id == 1 & treatment == "Placebo") %>% 
  ggplot(aes(area, amb_pm_24hr)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
                 position=position_nudge(x = -0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
                 size=7, position=position_nudge(x = -0.35, y = 0)) +
    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
                 position=position_nudge(x = 0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
                 size=7, position=position_nudge(x = 0.35, y = 0)) +
    theme_minimal() +
    labs(y = expression(paste("Ambient PM "[2.5], " (", mu, g/m^3, ")"))) +
    labs(title = "Ambient PM levels during indoor PM sampling",
          subtitle = "Data from closest EPA monitor to each home's zipcode") +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          title = element_text(size = 16, colour = "black"))
boxplots_amb_pm
```

```{r}
boxplots_amb_temp <- exposure_analysis_data_medium %>% 
  #filter(winter_id == 1 & treatment == "Placebo") %>% 
  ggplot(aes(area, mean_temp)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
                 position=position_nudge(x = -0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
                 size=7, position=position_nudge(x = -0.35, y = 0)) +
    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
                 position=position_nudge(x = 0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
                 size=7, position=position_nudge(x = 0.35, y = 0)) +
    theme_minimal() +
    labs(y = "Degrees (F)") +
    labs(title = "Ambient mean temperatures during indoor PM sampling",
          subtitle = "Data from closest site to each home's zipcode") +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          title = element_text(size = 16, colour = "black"))
boxplots_amb_temp
```

```{r}
wmt_map <- map_data("state") %>% 
  filter(region == "montana")

nn_map <- map_data("state") %>% 
  filter(region == "arizona")

ak_map <- map_data("world") %>% 
  filter(subregion == "Alaska") %>% 
  filter(long < 0)

wmt_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "WMT") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_pm = mean(amb_pm_24hr, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "azure", size = 1.2) +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_pm, size = mean_amb_pm), 
               shape = 21, color = "black") +
    geom_point(aes(long_pm, lat_pm), shape = 24, size = 3, fill = "yellow",
               color = "black") +
    theme_void() +
    labs(title = "WMT mean ambient PM at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangles = EPA monitors",
         fill = "Ambient PM", size = "Ambient PM") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 8)
wmt_plot
```

```{r}
nn_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "NN") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_pm = mean(amb_pm_24hr, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "azure", size = 1.2) +
    geom_point(aes(long_pm, lat_pm), shape = 24, size = 3, fill = "yellow",
               color = "black") +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_pm, size = mean_amb_pm), 
               shape = 21, color = "black") +
    theme_void() +
    labs(title = "NN mean ambient PM at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangles = EPA monitors",
         fill = "Ambient PM", size = "Ambient PM") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 1.5)
nn_plot
```

```{r}
ak_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "AK") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_pm = mean(amb_pm_24hr, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "azure", size = 1.2) +
    geom_point(aes(long_pm, lat_pm), shape = 24, size = 5, fill = "yellow",
               color = "black") +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_pm, size = mean_amb_pm), 
               shape = 21, color = "black") +
    theme_void() +
    labs(title = "AK mean ambient PM at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangle = EPA monitor",
         fill = "Ambient PM", size = "Ambient PM") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 6)
ak_plot
```

```{r}
wmt_map <- map_data("state") %>% 
  filter(region == "montana")

nn_map <- map_data("state") %>% 
  filter(region == "arizona")

ak_map <- map_data("world") %>% 
  filter(subregion == "Alaska") %>% 
  filter(long < 0)

wmt_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "WMT") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_temp = mean(mean_temp, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "azure", size = 1.2) +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_temp, size = mean_amb_temp), 
               shape = 21, color = "black") +
    geom_point(aes(long_temp, lat_temp), shape = 24, size = 3, fill = "yellow",
               color = "black") +
    theme_void() +
    labs(title = "WMT mean ambient Temp at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangles = weather stations",
         fill = "Ambient temp", size = "Ambient temp") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 20)
wmt_plot
```

```{r}
nn_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "NN") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_temp = mean(mean_temp, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "azure", size = 1.2) +
    geom_point(aes(long_temp, lat_temp), shape = 24, size = 3, fill = "yellow",
               color = "black") +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_temp, size = mean_amb_temp), 
               shape = 21, color = "black") +
    theme_void() +
    labs(title = "NN mean ambient Temp at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangles = weather stations",
         fill = "Ambient temp", size = "Ambient temp") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 45)
nn_plot
```

```{r}
ak_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "AK") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_temp = mean(mean_temp, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "azure", size = 1.2) +
    geom_point(aes(long_temp, lat_temp), shape = 24, size = 5, fill = "yellow",
               color = "black") +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_temp, size = mean_amb_temp), 
               shape = 21, color = "black") +
    theme_void() +
    labs(title = "AK mean ambient Temp at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangles = weather stations",
         fill = "Ambient temp", size = "Ambient temp") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 30)
ak_plot
```