### 2020-04-09 lrti_tracking.Rmd
exposure and lrti analysis updates

---
title: "Kids LRTI tracking"
author: "Ethan Walker"
date: "Started 6 April 2020, Updated 9 April 2020"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
```

```{r}
health_exposure_data_sampling_day <- read_rds("Output/health_exposure_data_sampling_day.rds")
```

# Diagnosing LRTI from the data

Is there a specific protocol Paul uses for diagnosing LRTI, or is it case-by-case
based on the given information for each child? We'll need to have some kind of
protocol/algorithm for publications, I would think.

Other questions:
- LRTI and ARI have been used interchangably. Our actual outcome is LRTI, yes?
  - Is there a clinical definition we/Paul have been following when diagnosing cases?
  - LRTI: bronchiolitis, pneumonia, croupe? Case definition up to Paul.
  - Primary diagnosis is based on med recs, but secondary outcomes could be from
    surveys. Also see how well survey data predicts Pauls diagnosis.
- How does Paul initially narrow down which cases to look at? Or does he look
  at each CSQ and med visit form?
- How far has the team made it (date/winter) in diagnosing LRTI? Winter 17/18?
  - Could touch base with Nick, but probably through winter 17/18.
- Has this process been done/started for NN/AK, or just WMT?
  - No. Surveys may be worse for these areas, but med rec data could be better
    through Indian Health Services.
  - At risk period could be difficult for these sites
- Where did the med_visit_ari variable come from?

Notes and specific documentation for steps to identify LRTI cases:
- Double check the filter method of "yes" for sick/still sick
- Ask Kathrene about med rec ARI vs med visit ARI and where they came from
- Was there a variable where parents called us to report sickness?
-

Next steps
- Ask Kathrene some questions from above
  - Have her upload final health data for NN/AK
- Push to get the final data input, QA/QC, uploaded to Box
- Go through winter 18/19 data line by line and flag cases that could be 
  respiratory.
  - Do this for all sites, and start on other sites for all years.
- Also flag other potential secondary outcomes that we could look at.

```{r}
# LRTI Diagnosis data
lrti_diagnosis_data <- health_exposure_data_sampling_day %>% 
  # select relevant health data
  select(area, home, home_winter_id, winter_id, child_id_num, 
         visit_date, med_visit_date, med_visit_referred,
         csq_date:csq_symptom_free_date,
         csq_see_dr, csq_dr_diagnosis,
         med_rec_visit_date:med_rec_comments,
         csq_miss_work:csq_comments,
         med_visit_ari:med_visit_mean_temp,
         lrti_number:lrti_professional_comments) %>% 
  # Currently only have this data for WMT - need to get new files and updates dataset
  filter(area == "WMT") %>% 
  # This is the most broad way of distinguishing if a child was sick
  # Careful when filtering here. Could still make cases that these vars didn't catch
  # Best to have a look through all rows, then use this to filter data down so it
  # is easier to work with to assess dates/at-risk times
  filter(csq_sick_2_weeks == "Yes" | csq_still_sick == "Yes") %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  ungroup()

#  filter(visit_date > "2018-06-01") # filter for newer data by date

# write_rds(lrti_diagnosis_data, "Output/lrti_diagnosis_data.rds")
# write_csv(lrti_diagnosis_data, "Output/lrti_diagnosis_data.csv", na = " ")
```

The filtering below narrows down the LRTI definition too much; we are missing cases
where children had LRTI but did not visit a healthcare facility and have a med rec.
If we only use these cases, it could bias the results.

There are instances of LRTI in our records that weren’t confirmed in the medical record. 
There are other instances where the medical record diagnosed LRTI, but our records 
overturned it. What is our criteria for diagnosing LRTI, where is it documented, 
and how can we be sure it is consistent in diagnosing LRTI?

```{r}
# Further filtering and re-ordering of variables
lrti_diagnosis_data_filtered <- lrti_diagnosis_data %>% 
  select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
         med_rec_visit_date:med_rec_comments, med_visit_lethargy:csq_comments,
         lrti_number:lrti_professional_comments) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, csq_collect) %>% 
  # If a home CSQ was done on same visit date as initial phone CSQ, filter for only the home CSQ
  distinct(csq_sick_2_weeks, .keep_all = TRUE) %>% 
  select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
         csq_first_symptom_date:lrti_professional_comments) %>% 
  # Were they referred (med visit) or did they see a Dr (CSQ)?
  filter(med_visit_referred == "Yes" | csq_see_dr == "Yes") %>% 
  # If yes, did we get medical records?
  filter(!is.na(med_rec_visit_date)) %>% 
  # If yes, was there an ARI diagnosis?
  filter(med_rec_ari == "ARI" | med_rec_ari == "Possible ARI") %>% 
  group_by(child_id_num) %>% 
  distinct(med_rec_visit_date, .keep_all = TRUE)
```

# Defining start date, end date, and at-risk period for each diagnosed LRTI case

First, I need to know (or make) exactly which variable is a final diagnosed case 
of LRTI in the data. Then, I can create a method in R to identify start/end/at-risk.
The code below is a start at this, using Nick's handoff documents/algorithm.

The initial attempt at having R select start dates did not match up well with
the cases in which Paul had already finalized the start/end dates. There seems
to be a good amount of professional opinion by assessing symptoms and medical
records. It may be best to filter a dataset down to only confirmed LRTI cases
and useful health/medical variables, then decide start/end/at-risk times 
on a case-by-case basis.

```{r}
# LRTI start date, end date, and at-risk period identification
lrti_date_data <- health_exposure_data_sampling_day %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         med_visit_date:med_visit_notes, chw_dr_relook:chw_ari_comments,
         csq_date:csq_comments, med_rec_visit_date:med_rec_comments,
         lrti_number:lrti_professional_comments) %>% 
  filter(area == "WMT") %>% 
  filter(csq_sick_2_weeks == "Yes") %>% 
  select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
         med_rec_visit_date:med_rec_comments, 
         med_visit_lethargy:lrti_professional_comments) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, csq_collect) %>% 
  distinct(csq_sick_2_weeks, .keep_all = TRUE) %>% 
  select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
         csq_first_symptom_date:lrti_professional_comments) %>% 
  mutate(trash_date = ymd("2099-01-01"),
         still_sick_date = if_else(csq_still_sick == "Yes", 
                                   csq_date, trash_date),
         med_rec_ari_date = if_else(!is.na(med_rec_visit_date), 
                                    med_rec_visit_date, trash_date),
         med_visit_ari_date = if_else(!is.na(med_visit_date), 
                                      med_visit_date, trash_date)) %>% 
  pivot_longer(cols = c("still_sick_date", "med_rec_ari_date", "med_visit_ari_date"),
               names_to = "lrti_visits", values_to = "lrti_dates") %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  mutate(lrti_start_date_new = if_else(!is.na(csq_first_symptom_date),
                                       csq_first_symptom_date, lrti_dates)) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         csq_first_symptom_date, lrti_start_date_new, lrti_visits, lrti_dates, trash_date,
         csq_sick_2_weeks, csq_still_sick, med_rec_ari, med_visit_ari,
         lrti_number:lrti_professional_comments) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, lrti_dates) %>% 
  distinct(visit_date, .keep_all = TRUE) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num) %>% 
  distinct(lrti_start_date_new, .keep_all = TRUE) %>% 
  distinct(child_id_num, lrti_number, .keep_all = TRUE) %>% 
  filter(!is.na(lrti_number)) %>% 
  select(child_id_num, home_winter_id, lrti_number, lrti_start_date_new:lrti_dates, 
         lrti_start_date_final, lrti_end_date_final, area:lrti_professional_comments)
```