### 2020-04-14 child_health_data_cleaning.Rmd
Update PM data

---
title: "Kids health and child data cleaning"
author: "Ethan Walker"
date: "Started 27 Feb 2020, Updated 14 April 2020"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_linked_ids <- read_rds(paste0(output_path, "Output/kids_linked_ids.rds"))
demographics_clean <- 
  read_rds(paste0(output_path, "Output/demographics_clean.rds"))
pm_clean <- read_rds(paste0(output_path, "Output/pm_clean.rds"))
```

################# At Home Log ###############

# Load and merge files from different locations
```{r}
wmt_athomelog_initial <- read_xlsx("Input/WMT/wmt_athomelog_20200227.xlsx") %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "WMT",
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

nn_athomelog_initial <- read_xlsx("Input/NN/nn_athomelog_20200227.xlsx") %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "NN",
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

ak_athomelog_initial <- read_xlsx("Input/AK/ak_athomelog_20200227.xlsx") %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "AK",
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

# Combine logs from all areas
kids_at_home_log <- rbind(wmt_athomelog_initial, nn_athomelog_initial,
                          ak_athomelog_initial) %>% 
  left_join(kids_linked_ids, c("area", "home_winter_id"))

# write_rds(kids_at_home_log, "Output/kids_at_home_log.rds")
```

# Format at-home-log and merge with PM data
```{r}
kids_at_home_log <- 
  read_rds(paste0(output_path, "Output/kids_at_home_log.rds")) %>% 
  arrange(area, home_winter_id, at_home_day)

kids_at_home_log_new <- kids_at_home_log %>% 
  mutate(sampling_day = at_home_day) %>% 
  group_by(area, home_winter_id, child_id_num, sampling_day) %>% 
  mutate("00:00:00" = 1,
         "01:00:00" = 1,
         "02:00:00" = 1,
         "03:00:00" = 1,
         "04:00:00" = 1,
         "05:00:00" = 1,
         "06:00:00" = 1,
         "07:00:00" = as.numeric(at_home_7) - 1,
         "08:00:00" = as.numeric(at_home_8) - 1,
         "09:00:00" = as.numeric(at_home_9) - 1,
         "10:00:00" = as.numeric(at_home_10) - 1,
         "11:00:00" = as.numeric(at_home_11) - 1,
         "12:00:00" = as.numeric(at_home_12) - 1,
         "13:00:00" = as.numeric(at_home_13) - 1,
         "14:00:00" = as.numeric(at_home_14) - 1,
         "15:00:00" = as.numeric(at_home_15) - 1,
         "16:00:00" = as.numeric(at_home_16) - 1,
         "17:00:00" = as.numeric(at_home_17) - 1,
         "18:00:00" = as.numeric(at_home_18) - 1,
         "19:00:00" = as.numeric(at_home_19) - 1,
         "20:00:00" = as.numeric(at_home_20) - 1,
         "21:00:00" = as.numeric(at_home_21) - 1,
         "22:00:00" = as.numeric(at_home_22) - 1,
         "23:00:00" = 1) %>% 
  pivot_longer(cols = c("00:00:00":"23:00:00"), names_to = "sample_time",
                        values_to = "at_home") %>% 
  mutate("00:00:00" = "home",
         "01:00:00" = "home",
         "02:00:00" = "home",
         "03:00:00" = "home",
         "04:00:00" = "home",
         "05:00:00" = "home",
         "06:00:00" = "home",
         "07:00:00" = location_7,
         "08:00:00" = location_8,
         "09:00:00" = location_9,
         "10:00:00" = location_10,
         "11:00:00" = location_11,
         "12:00:00" = location_12,
         "13:00:00" = location_13,
         "14:00:00" = location_14,
         "15:00:00" = location_15,
         "16:00:00" = location_16,
         "17:00:00" = location_17,
         "18:00:00" = location_18,
         "19:00:00" = location_19,
         "20:00:00" = location_20,
         "21:00:00" = location_21,
         "22:00:00" = location_22,
         "23:00:00" = "home") %>% 
  select(area, home, home_winter_id, child_id_num, winter_id, treatment,
         sampling_day, sample_time, at_home, "00:00:00":"23:00:00") %>% 
  group_by(area, home_winter_id, child_id_num, sampling_day, sample_time) %>% 
  pivot_longer(cols = c("00:00:00":"23:00:00"), names_to = "location_time",
                        values_to = "at_home_location") %>% 
  mutate(filter_var = if_else(sample_time == location_time, 1, 0)) %>% 
  filter(filter_var == 1) %>%  
  arrange(area, home_winter_id, child_id_num, sampling_day, sample_time) %>% 
  select(area, home, home_winter_id, child_id_num, winter_id, treatment,
         sampling_day, sample_time, at_home, at_home_location)  %>% 
  group_by(area, home_winter_id, child_id_num, sampling_day) %>% 
  mutate(at_home = if_else(sample_time == "06:00:00", lead(at_home), at_home),
         at_home = if_else(sample_time == "05:00:00", lead(at_home), at_home),
         at_home = if_else(sample_time == "04:00:00", lead(at_home), at_home),
         at_home = if_else(sample_time == "03:00:00", lead(at_home), at_home),
         at_home = if_else(sample_time == "02:00:00", lead(at_home), at_home),
         at_home = if_else(sample_time == "01:00:00", lead(at_home), at_home),
         at_home = if_else(sample_time == "00:00:00", lead(at_home), at_home),
         at_home = if_else(sample_time == "23:00:00", lag(at_home), at_home),
         at_home_location = if_else(sample_time == "06:00:00", 
                                    lead(at_home_location), at_home_location),
         at_home_location = if_else(sample_time == "05:00:00", 
                                    lead(at_home_location), at_home_location),
         at_home_location = if_else(sample_time == "04:00:00", 
                                    lead(at_home_location), at_home_location),
         at_home_location = if_else(sample_time == "03:00:00", 
                                    lead(at_home_location), at_home_location),
         at_home_location = if_else(sample_time == "02:00:00", 
                                    lead(at_home_location), at_home_location),
         at_home_location = if_else(sample_time == "01:00:00", 
                                    lead(at_home_location), at_home_location),
         at_home_location = if_else(sample_time == "00:00:00", 
                                    lead(at_home_location), at_home_location),
         at_home_location = if_else(sample_time == "23:00:00", 
                                    lag(at_home_location), at_home_location),
         at_home_location = if_else(at_home == 1, "home", at_home_location)) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, child_id_num, sampling_day, sample_time)

at_home_pm_merged <- pm_clean %>% 
  group_by(area, home) %>% 
  mutate(pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE),
         pm_datetime_merge = round_date(pm_datetime_new, unit = "hour"),
         pm_obs_sampling_day = n()) %>% 
  separate(pm_datetime_merge, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home, sampling_day, sample_time) %>% 
  mutate(pm_mean_hour = mean(pm, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  distinct(sample_time, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(sampling_day = as.character(sampling_day)) %>% 
  select(area, home, home_winter_id, winter_id, treatment, sampling_day, 
         sample_date, sample_time, pm_mean_hour, pm_mean_sampling_day, 
         pm_mean_sampling_period, day_of_week, pm_obs_sampling_day, 
         pm_spike_1000, pm_spike_2000, pm_spike_5000) %>% 
  left_join(kids_at_home_log_new, 
            by = c("area", "home", "home_winter_id", "winter_id", "treatment",
                   "sampling_day", "sample_time")) %>% 
  group_by(area, home, home_winter_id, child_id_num, at_home) %>% 
  mutate(pm_at_home_sampling_period = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  group_by(area, home, home_winter_id, child_id_num, sampling_day, at_home) %>% 
  mutate(pm_at_home_sampling_day = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, child_id_num, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  distinct(at_home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  filter(!is.na(at_home)) %>% 
  arrange(area, home_winter_id, child_id_num, at_home, sampling_day)

#write_rds(at_home_pm_merged, paste0(output_path, "Output/at_home_pm_merged.rds"))
#write_csv(at_home_pm_merged, paste0(output_path, "Output/at_home_pm_merged.csv"))

at_home_pm_merged <- 
  read_rds(paste0(output_path, "Output/at_home_pm_merged.rds"))
```

################# Medical Visit ###############

# Load and merge files from different locations
```{r}
wmt_medicalvisit_initial <- read_xlsx("Input/WMT/wmt_medicalvisit_20200227.xlsx") %>% 
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

nn_medicalvisit_initial <- read_xlsx("Input/NN/nn_medicalvisit_20200227.xlsx") %>%
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "NN",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

ak_medicalvisit_initial <- read_xlsx("Input/AK/ak_medicalvisit_20200227.xlsx") %>% 
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "AK",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

# Combine logs from all areas
kids_med_visit <- rbind(wmt_medicalvisit_initial, nn_medicalvisit_initial,
                          ak_medicalvisit_initial) %>% 
  left_join(kids_linked_ids, c("area", "home_winter_id"))

# write_rds(kids_med_visit, "Output/kids_med_visit.rds")

kids_med_visit <- read_rds("Output/kids_med_visit.rds")

save_var_names <- data.frame(t(kids_med_visit)) %>% 
  rownames_to_column() %>% 
  select(rowname)

write_csv(save_var_names, "Output/medvisit_var_names.csv")
```

################# Birth Infancy ###############

# Load and merge files from different locations
```{r}
wmt_birthinfancy_initial <- read_xlsx("Input/WMT/wmt_birthinfancy_20200227.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         gender = as.character(Gender),
         bw_lbs = as.numeric(BirthweightLbs),
         bw_oz = as.numeric(BirthWeightOz),
         birth_weight_oz = as.numeric((bw_lbs*16)+bw_oz),
         mom_age = as.numeric(AgeOfMom),
         breastfeed_current = as.character(CurrentlyBreastfeeding),
         breastfeed_past = as.character(DidBreastfeed),
         bf_years = as.numeric(HowLongYears),
         bf_months = as.numeric(HowLongMonths),
         breastfeed_months = as.numeric((bf_years*12)+bf_months),
         flu_shot = as.character(FluShot),
         flu_shot_planned = as.character(FluShotPlanned),
         asthma_meds = as.character(AsthmaMeds),
         child_school = as.character(Attend),
         school_hours = as.numeric(HoursDay),
         school_days = as.numeric(DaysWeek),
         school_hours_week = as.numeric(school_hours*school_days),
         sibling_school = as.character(SibSchool)) %>% 
  select(home_winter_id:sibling_school)

# Waiting on files from NN and AK

# Combine logs from all areas
kids_birth_infancy <- wmt_birthinfancy_initial %>% 
  left_join(kids_linked_ids, c("area", "home_winter_id"))

# write_rds(kids_birth_infancy, "Output/kids_birth_infancy.rds")

kids_birth_infancy <- read_rds("Output/kids_birth_infancy.rds")

save_var_names <- data.frame(t(wmt_birthinfancy_initial)) %>% 
  rownames_to_column() %>% 
  select(rowname)

write_csv(save_var_names, "Output/birthinfancy_var_names.csv")
```

################# Child Home Winter  ###############

# Load and merge files from different locations
```{r}
wmt_childhomewinter_initial <- read_xlsx("Input/WMT/wmt_childhomewinter_20200227.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         home_id = as.character(HomeID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         chw_dr_relook = DrRelook,
         chw_ari_status = as.character(Status),
         chw_symptom_free_date = ymd(SymptomFreeDate),
         chw_csq_duedate = ymd(CSQDueDate),
         chw_csq_date = ymd(CSQDate),
         chw_csq_time = as.character(CSQTime),
         chw_dr_comments = as.character(DrComments),
         chw_dr_history = as.character(DrHistory),
         chw_notify_parents = as.character(NotifyParents),
         chw_notes_to_parents = as.character(NotesToParents),
         chw_any_ari = as.character(AnyARI),
         chw_ari_comments = as.character(WinterARIComments)) %>% 
  select(home_winter_id:chw_ari_comments)

# Waiting on files from NN and AK

# Combine logs from all areas
child_home_winter <- wmt_childhomewinter_initial %>% 
  left_join(kids_linked_ids, c("area", "home_winter_id"))

# write_rds(child_home_winter, "Output/child_home_winter.rds")

child_home_winter <- read_rds("Output/child_home_winter.rds")

```

################# CSQ - child symptom questionnaire  ###############

# Load and merge files from different locations
```{r}
wmt_csq_initial <- read_xlsx("Input/WMT/wmt_csq_20200227.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         csq_date = ymd(Date),
         csq_collect = as.character(HowCollected),
         csq_sick_2_weeks = as.character(Illin2Weeks),
         csq_still_sick = as.character(StillSick),
         csq_first_symptom_date = ymd(FirstSymptomDate),
         csq_symptom_free_date = ymd(SymptomFreeDate),
         csq_miss_work = as.character(MissWork),
         csq_miss_work_days = as.character(DaysMissed),
         csq_see_dr = as.character(SeeDr),
         csq_dr_diagnosis = as.character(DrDiagnosis),
         csq_new_meds = as.character(NewMeds),
         csq_fever_2weeks = as.character(Fever),
         csq_temp_2weeks = as.numeric(Temperature),
         csq_rapid_breathing_2weeks = as.character(RapidBreathing),
         csq_difficult_breathing_2weeks = as.character(DifficultBreathing),
         csq_noisy_breathing_2weeks = as.character(NoisyBreathing),
         csq_runny_nose_2weeks = as.character(RunnyNose),
         csq_cough_2weeks = as.character(Coughing),
         csq_wheeze_2weeks = as.character(Wheezing),
         csq_fever_current = as.character(FeverStill),
         csq_rapid_breathing_current = as.character(RapidBreathingStill),
         csq_difficult_breathing_current = as.character(DifficultBreathingStill),
         csq_noisy_breathing_current = as.character(NoisyBreathingStill),
         csq_runny_nose_current = as.character(RunnyNoseStill),
         csq_cough_current = as.character(CoughingStill),
         csq_wheeze_current = as.character(WheezingStill),
         csq_comments = as.character(Comments)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -99, -99.9, 999, 0)) %>% 
  select(home_winter_id:csq_comments)

# Waiting on files from NN and AK

# Combine logs from all areas
kids_csq <- wmt_csq_initial %>% 
  left_join(kids_linked_ids, c("area", "home_winter_id"))

# write_rds(kids_csq, "Output/kids_csq.rds")

kids_csq <- read_rds("Output/kids_csq.rds")

save_var_names <- data.frame(t(wmt_csq_initial)) %>% 
  rownames_to_column() %>% 
  select(rowname)

write_csv(save_var_names, "Output/csq_var_names.csv")
```

################# Medical Records ###############

# Load and merge files from different locations
```{r}
wmt_medicalrecords_initial <- read_xlsx("Input/WMT/wmt_medicalrecords_20200227.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         visit_date = ymd(VisitDate),
         med_rec_visit_date = ymd(VisitDate),
         med_rec_dr_diagnosis = as.character(DrDiag),
         med_rec_ari = as.character(ARI),
         med_rec_comments = as.character(DrComments)) %>% 
  select(home_winter_id:med_rec_comments)

# Waiting on files from NN and AK

# Combine logs from all areas
kids_medical_records <- wmt_medicalrecords_initial %>% 
  left_join(kids_linked_ids, c("area", "home_winter_id"))

# write_rds(kids_medical_records, "Output/kids_medical_records.rds")

kids_medical_records <- read_rds("Output/kids_medical_records.rds")
```

################# Confirmed LRTI ###############

# Load and merge files from different locations
```{r}
wmt_ari_initial <- read_xlsx("Input/WMT/wmt_ari_20200408.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         lrti_number = as.character(ARINum),
         lrti_basis = as.character(ARIBasis),
         lrti_start_date_final = ymd(ProfessionalStartDate),
         lrti_end_date_final = ymd(ProfessionalEndDate),
         lrti_professional_comments = as.character(ProfessionalComments)) %>% 
  select(home_winter_id:lrti_professional_comments)

# Waiting on files from NN and AK

# Combine logs from all areas
kids_lrti <- wmt_ari_initial %>% 
  left_join(kids_linked_ids, c("area", "home_winter_id"))

# write_rds(kids_lrti, "Output/kids_lrti.rds")

kids_lrti <- read_rds("Output/kids_lrti.rds")

```