### 2020-04-14 merge_health_datasets.Rmd

---
title: "KidsAIR: merge health datasets"
author: "Ethan Walker"
date: "Started 27 March 2020, Updated 14 April 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(lme4)
library(lmerTest)
```

# Load individual datasets
```{r}
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

at_home_pm_merged <- 
  read_rds(paste0(output_path, "Output/at_home_pm_merged.rds"))
kids_med_visit <- 
  read_rds(paste0(output_path, "Output/kids_med_visit.rds")) %>% 
  mutate(visit_date = med_visit_date) 
kids_birth_infancy <- 
  read_rds(paste0(output_path, "Output/kids_birth_infancy.rds")) %>% 
  select(-home_id)
child_home_winter <- 
  read_rds(paste0(output_path, "Output/child_home_winter.rds")) %>% 
  select(-home_id.x, -home_id.y, -chw_symptom_free_date)
kids_csq <- 
  read_rds(paste0(output_path, "Output/kids_csq.rds"))
kids_lrti <- 
  read_rds(paste0(output_path, "Output/kids_lrti.rds")) %>% 
  select(-home_id)
kids_medical_records <- 
  read_rds(paste0(output_path, "Output/kids_medical_records.rds")) %>% 
  select(-visit_date)
```

# Initial data merge - long format
```{r}
health_merged1 <- kids_csq %>% 
  select(-home_id) %>% 
  mutate(visit_date = csq_date) %>% 
  full_join(kids_med_visit, by = c("area", "home", "home_winter_id", "treatment",
                                   "winter_id", "child_id_num", "visit_date")) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         treatment, csq_date:med_visit_mean_temp) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  ungroup()

health_merged2 <- health_merged1 %>% 
  left_join(child_home_winter, by = c("area", "home", "home_winter_id", "treatment",
                                      "winter_id", "child_id_num")) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  ungroup()

health_merged3 <- health_merged2 %>% 
  left_join(kids_birth_infancy, by = c("area", "home", "home_winter_id", "treatment",
                                      "winter_id", "child_id_num")) %>% 
  mutate(sample_date = as.character(visit_date)) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  ungroup()

health_merged4 <- health_merged3 %>% 
  full_join(at_home_pm_merged, by = c("area", "home", "home_winter_id", "winter_id", 
                                      "child_id_num", "sample_date", "treatment")) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, sample_date) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         treatment, sample_date, sampling_day, at_home_location, 
         pm_at_home_sampling_day, pm_at_home_sampling_period, csq_date:at_home) %>% 
  ungroup()

health_merged5 <- health_merged4 %>% 
  left_join(kids_medical_records, by = c("area", "home", "home_winter_id", "winter_id", 
                                      "child_id_num", "treatment")) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, sample_date) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         treatment, sample_date, sampling_day, at_home_location, 
         pm_at_home_sampling_day, pm_at_home_sampling_period, csq_date:med_rec_comments) %>% 
  ungroup()

health_merged6 <- health_merged5 %>% 
  left_join(kids_lrti, by = c("area", "home", "home_winter_id", "winter_id", 
                                      "child_id_num", "treatment")) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, sample_date) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         treatment, sample_date, sampling_day, at_home_location, 
         pm_at_home_sampling_day, pm_at_home_sampling_period, csq_date:lrti_professional_comments) %>% 
  ungroup()

#write_rds(health_merged6, paste0(output_path, "Output/health_data_merged.rds"))
#write_csv(health_merged6, paste0(output_path, "Output/health_data_merged.csv"))
```

# Merge full health data with exposure data
```{r}
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

exposure_analysis_data_medium <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))

health_data_merged <- 
  read_rds(paste0(output_path, "Output/health_data_merged.rds"))

health_exposure_data_sampling_day <- health_data_merged %>% 
  select(-pm_mean_sampling_day, -pm_mean_sampling_period, day_of_week) %>% 
  mutate(winter_id = as.character(winter_id)) %>% 
  left_join(exposure_analysis_data_medium,
            by = c("area", "home", "home_winter_id", "winter_id", "treatment",
                   "child_id_num", "sampling_day"))

#write_rds(health_exposure_data_sampling_day, paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
#write_csv(health_exposure_data_sampling_day, paste0(output_path, "Output/health_exposure_data_sampling_day.csv"))

health_exposure_data_sampling_day <- 
  read_rds(paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))

save_var_names <- data.frame(t(health_exposure_data_sampling_day)) %>% 
  rownames_to_column() %>% 
  select(rowname)

#write_csv(save_var_names, paste0(output_path, "Output/kids_data_var_names.csv"))
```