### 2020-04-16 exposure_initial_analysis.Rmd
exposure analysis update

---
title: "KidsAIR: initial exposure analysis"
author: "Ethan Walker"
date: "Started 2 Feb 2020, Updated 13 April 2020"
output: pdf_document
header-includes:
    - \usepackage[labelformat=empty]{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
```

# Load individual datasets
```{r}
input_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/Input/")
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
input_file <- c("exposure_analysis_data_long_new.rds")
output_file <- c("exposure_analysis_data_long_new.rds")

exposure_analysis_data_long <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))

exposure_analysis_data_medium <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))

exposure_analysis_data_short <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_short.rds"))

sums_rolling_pm_data <- 
  read_rds(paste0(output_path, "Output/sums_rolling_pm_data.rds"))

health_exposure_data_sampling_day <- 
  read_rds(paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
```

# Demographics
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  ungroup()

# Variables to use:
## education_3level, income_3level, race_parent_2level, 
## total_residents_2level, residents_under5_2level, resident_smokes

table(analysis_data$residents_smoke)
median(analysis_data$home_furry, na.rm = TRUE)

pm_demographics_results <- lm(log(pm_mean_sampling_period) ~ education_3level, 
                              data = analysis_data)

summary(pm_demographics_results)
glance(pm_demographics_results)

tidy(pm_demographics_results) %>% 
    filter(term != '(Intercept)') %>% 
    mutate(estimate = round(estimate, digits = 2),
           'p-value' = round(p.value, digits = 2),
           term = gsub("education_3level", "Education:  ", term)) %>% 
    select(term, estimate, 'p-value')

plot(pm_demographics_results)
```

# Home characteristics
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## home_type_2level, home_year_built_2level, home_sqft_2level, home_floors_2level
## home_bedrooms_2level, home_windows_2level, home_pets_2level, home_furry_2level
## secondary_heat_3level, stove_age_3level, stove_cert, chimney_clean_3level,
## home_damp, home_mold, home_fireplace, home_furnace, home_insert

table(analysis_data$home_insert)
median(analysis_data$home_furry, na.rm = TRUE)

pm_home_char_results <- lm(log(pm_mean_sampling_period) ~ home_insert, 
                           data = analysis_data)
summary(pm_home_char_results)
tidy(pm_home_char_results)
plot(pm_home_char_results)
```

# Stove use
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## burn_level_3level, wood_collect_2level, wood_collect_method_2level,
## wood_cords

table(analysis_data$burn_level_3level)
median(analysis_data$wood_cords, na.rm = TRUE)

pm_stove_use_results <- lm(log(pm_mean_sampling_period) ~ wood_collect_method_2level, 
                           data = analysis_data)
summary(pm_stove_use_results)
tidy(pm_stove_use_results)
plot(pm_stove_use_results)
pm_stove_use_plot <- pm_stove_use %>% 
  ggplot() + 
    geom_boxplot(aes(burn_level, mean_pm))
pm_stove_use_plot
```

# Stove grades
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## stove_grade_3level

table(analysis_data$stove_grade_3level)

pm_stove_grades_results <- lm(log(pm_mean_sampling_period) ~ stove_grade_3level, 
                              data = analysis_data)
summary(pm_stove_grades_results)
tidy(pm_stove_grades_results)
anova(pm_stove_grades_results)
plot(pm_stove_grades_results)
pm_stove_grades_plot <- pm_stove_grades %>% 
  ggplot() + 
    geom_boxplot(aes(stove_grade, mean_pm))
pm_stove_grades_plot
```

```{r}
analysis_function <- function(covar, label, 
                              dataset = exposure_analysis_data_short){

  analysis_data <- dataset %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  rename(analysis_covar = covar) %>% 
  ungroup() %>% 
  select(pm_mean_sampling_period, analysis_covar)

  analysis_results <- lm(log(pm_mean_sampling_period) ~ analysis_covar, analysis_data)

  tidy_results <<- tidy(analysis_results) %>% 
    filter(term != '(Intercept)') %>% 
    mutate(estimate = round(estimate, digits = 2),
           'p-value' = round(p.value, digits = 2),
           term = gsub("analysis_covar", paste0(label, ": "), term)) %>% 
    select(term, estimate, 'p-value')

}

d1 <- analysis_function("education_3level", "Education")
d2 <- analysis_function("income_3level", "Income")
d3 <- analysis_function("race_parent_2level", "Parent race")
d4 <- analysis_function("total_residents_2level", "Total residents")
d5 <- analysis_function("residents_under5_2level", "Residents under 5 y/o")
d6 <- analysis_function("residents_smoke", "Residents smoke")

d7 <- analysis_function("home_type_2level", "Home type")
d8 <- analysis_function("home_year_built_2level", "Home age")
d9 <- analysis_function("home_sqft_2level", "home sqft")
d10 <- analysis_function("home_floors_2level", "Home floors")
d11 <- analysis_function("home_bedrooms_2level", "Home bedrooms")
d12 <- analysis_function("home_windows_2level", "Home windows")
d13 <- analysis_function("home_pets_2level", "Pets - total")
d14 <- analysis_function("home_furry_2level", "Pets - furry")
d15 <- analysis_function("secondary_heat_3level", "Secondary heating")
d16 <- analysis_function("stove_age_3level", "Stove age")
d17 <- analysis_function("stove_cert", "Stove EPA certified")
d18 <- analysis_function("chimney_clean_3level", "Chimney cleaned")
d19 <- analysis_function("home_damp", "Home - damp")
d20 <- analysis_function("home_mold", "Home - mold")
d21 <- analysis_function("home_fireplace", "Home fireplace")
d22 <- analysis_function("home_furnace", "Home furnace")

d23 <- analysis_function("burn_level_3level", "Burn level")
d24 <- analysis_function("wood_collect_2level", "Wood collect time")
d25 <- analysis_function("wood_collect_method_2level", "Wood collect method")
d26 <- analysis_function("wood_cords", "Wood cords")

table_data <- rbind(d1, d2, d3, d4, d5, d6,
                    d7, d8, d9, d10, d11, d12, d13, d14, d15, d16, d17, d18,
                    d19, d20, d21, d22,
                    d23, d24, d25, d26)

kable(table_data, 
      caption = "Indoor PM2.5 (log) and covariate associations") %>% 
   kable_styling() %>% 
   row_spec(0, bold = TRUE) %>% 
   pack_rows("Demographic variables", 1, 8) %>%
   pack_rows("Home characteristics", 9, 31) %>%
   pack_rows("Stove use", 32, 36) %>%
   footnote(c("PM2.5 = fine particulate matter"))
```

# Home activity
```{r}
analysis_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## home_act_gas, home_act_elect, home_act_propane, home_act_oil, home_act_other,
## home_act_smoke, home_act_incense, home_act_candle, home_act_windows,
## home_act_door, home_act_sweep

table(analysis_data$home_act_sweep)

pm_home_act_results <- lmer(log(pm_mean_sampling_day) ~ home_act_sweep + 
                              (1 | home), data = analysis_data)
summary(pm_home_act_results)
tidy(pm_home_act_results)
plot(pm_home_act_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_home_act_results), main = "QQ Plot")
```

# Wood moisture content
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE))

# Variables to use:
## moisture_closest, moisture_closest_2level
## mean_moisture_winter, moisture_winter_2level

table(analysis_data$moisture_closest_2level)
median(analysis_data$mean_moisture_winter, na.rm = TRUE)

pm_moisture_results <- lm(log(pm_mean_sampling_period) ~ 
                            moisture_winter_2level, data = analysis_data)
summary(pm_moisture_results)
plot(pm_moisture_results)
```

# Indoor temp/RH
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE))

# Variables to use:
## rh_min_5, rh_max_5, temp_min_5, temp_max_5
## temp_max_2level, temp_min_2level,rh_max_2level ,rh_min_2level

table(analysis_data$temp_max_2level)
median(analysis_data$rh_indoor_min, na.rm = TRUE)

pm_temp_rh_results <- lm(log(pm_mean_sampling_period) ~ temp_max_2level, 
                         data = analysis_data)
summary(pm_temp_rh_results)
tidy(pm_temp_rh_results)
plot(pm_temp_rh_results)
```

# SUMs - by sampling day
```{r}
analysis_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  mutate(sums_mean_sampling_day_5 = sums_mean_sampling_day/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  ungroup()

# Variables to use:
## sums_mean_sampling_day_5, temp_perc_25_2level
## temp_perc_27_2level, temp_perc_30_2level

table(analysis_data$temp_perc_30_2level)
median(analysis_data$temp_perc_30, na.rm = TRUE)

pm_sums_results <- lmer(log(pm_mean_sampling_day) ~ temp_perc_30_2level + 
                              (1 | home), data = analysis_data)
pm_sums_results <- lmer(pm_mean_sampling_day ~ temp_perc_30_2level + 
                              (1 | home), data = analysis_data)
summary(pm_sums_results)
tidy(pm_sums_results)
plot(pm_sums_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_results), main = "QQ Plot")
```

# SUMs - by sampling period
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  ungroup()

# Variables to use:
## sums_mean_sampling_period_5, temp_perc_25_2level
## temp_perc_27_2level, temp_perc_30_2level

table(analysis_data$temp_perc_25_2level)
median(analysis_data$temp_perc_30, na.rm = TRUE)

pm_sums_results <- lm(log(pm_mean_sampling_period) ~ sums_mean_sampling_period_5, 
                      data = analysis_data)
pm_sums_results <- lm(pm_mean_sampling_period ~ sums_mean_sampling_period_5, 
                      data = analysis_data)
summary(pm_sums_results)
tidy(pm_sums_results)
plot(pm_sums_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_results), main = "QQ Plot")
```

# SUMs - mean PM per SUMs interval
```{r}
analysis_data <- sums_rolling_pm_data %>% 
  mutate(temp_c_sums_5 = temp_c_sums/5)

cor.test(analysis_data$temp_c_sums, analysis_data$pm_rolling_20, method = "spearman")
pm_sums_model <- lmer(log(pm_rolling_20) ~ temp_c_sums_5 + (1 | home), 
                      data = analysis_data)
pm_sums_model <- lmer(pm_rolling_20 ~ temp_c_sums_5 + (1 | home), 
                      data = analysis_data)

summary(pm_sums_model)
tidy(pm_sums_model)
exp((0.1496794)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

pm_sums_plot <- analysis_data %>% 
  ggplot() + 
    geom_point(aes(log(pm_rolling_20), temp_c_sums)) +
    geom_smooth(aes(log(pm_rolling_20), temp_c_sums)) +
    theme_minimal()
pm_sums_plot
```

# SUMs - events per sampling day 
```{r}
sums_temp_change <- sums_rolling_pm_data %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home, sampling_day) %>% 
  distinct(datetime_sums, .keep_all = TRUE) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(time_diff = (lead(datetime_sums) - datetime_sums)) %>% 
  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
  replace_with_na(replace = list(lead_temp = 999)) %>% 
  mutate(temp_diff = lead_temp - temp_c_sums,
         # change temp in the following line
         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
                              temp_diff_check),
         heat_event = if_else(is.na(temp_diff_check), 1,
                              heat_event)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(sums_events = sum(heat_event, na.rm = TRUE),
         sums_events_cat = cut(sums_events, breaks = c(0, 2, 50),
                           labels = c("<2", "2+"),
                                 right = FALSE)) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup() %>% 
  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
         time_diff:sums_events_cat) %>% 
  arrange(area, home, sample_date)

median(sums_temp_change$sums_events, na.rm = TRUE)
table(sums_temp_change$sums_events_cat)

pm_sums_model <- lmer(pm_mean_sampling_day ~ sums_events_cat + 
                        (1 | home), data = sums_temp_change)
pm_sums_model <- lmer(log(pm_mean_sampling_day) ~ sums_events_cat + 
                        (1 | home), data = sums_temp_change)

summary(pm_sums_model)
tidy(pm_sums_model)
#exp((0.146291)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

pm_sums_plot <- sums_temp_change %>% 
  filter(!is.na(sums_events)) %>% 
  ggplot() + 
    geom_boxplot(aes(sums_events_cat, log(pm_mean_sampling_day))) +
    theme_minimal()
pm_sums_plot
```

# Ambient temp/PM - by sampling date
```{r}
analysis_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(area == "AK") %>% 
  filter(!is.na(sampling_day)) %>% 
  filter(!is.na(sample_date)) %>%
  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
         zip:mean_temp_roll_4day) %>% 
  mutate(amb_pm_5 = amb_pm_24hr/5,
         mean_temp_5 = mean_temp/5,
         mean_temp_cat = cut(mean_temp, breaks = c(0, 30, 150),
                           labels = c("<30", "30+"),
                                 right = FALSE),
         amb_pm_cat = cut(amb_pm_24hr, breaks = c(0, 7, 100),
                           labels = c("<7", "7+"),
                                 right = FALSE))

median(analysis_data$amb_pm_24hr, na.rm = TRUE)
table(analysis_data$amb_pm_cat)

# Variables to use:
## amb_pm_5, amb_pm_cat,
## amb_pm_24hr, amb_pm_24hr_lag1, amb_pm_24hr_lag2, amb_pm_24hr_lag3
## amb_pm_roll_mean_2day, amb_pm_roll_mean_3day, amb_pm_roll_mean_4day
## mean_temp_5, mean_temp_cat,
## mean_temp, mean_temp_lag1, mean_temp_lag2, mean_temp_lag3
## mean_temp_roll_2day, mean_temp_roll_3day, mean_temp_roll_4day

cor.test(analysis_data$mean_temp, analysis_data$pm_mean_daily, method = "spearman")
pm_ambient_model <- lmer(log(pm_mean_daily) ~ amb_pm_cat + (1 | home), 
                      data = analysis_data)
pm_ambient_model <- lmer(pm_mean_daily ~ amb_pm_cat + (1 | home), 
                      data = analysis_data)

#summary(pm_ambient_model)
tidy(pm_ambient_model)
exp((-0.06723228)-1)*100
plot(pm_ambient_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_ambient_model), main = "QQ Plot")

pm_sums_plot <- analysis_data %>% 
  ggplot() + 
    geom_point(aes(log(pm_mean_daily), mean_temp)) +
    geom_smooth(aes(log(pm_mean_daily), mean_temp)) +
    theme_minimal()
pm_sums_plot
```

# Run tests for multicollinearity between different covariates
Goal is to check for vars that have a lot of overlap and eliminate some from
consideration to include in models

First, use `mctest` package to check for multicollinearity among similar variables
https://cran.r-project.org/web/packages/mctest/mctest.pdf

Notes on multicollinearity:

https://datascienceplus.com/multicollinearity-in-r/ 
https://www.theanalysisfactor.com/eight-ways-to-detect-multicollinearity/

```{r}
# Filter data for specific variables; select pm variable
pm_mean_sampling_period <- exposure_analysis_data_short %>% 
  filter(winter_id == 1) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  select(pm_mean_sampling_period, income_3level, residents_smoke) %>% 
  mutate_all(as.numeric) %>% 
  na.exclude() %>% 
  select(pm_mean_sampling_period)

# Filter data for specific variables; select predictor variables
other_vars <- exposure_analysis_data_short %>% 
  filter(winter_id == 1) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  select(pm_mean_sampling_period, income_3level, residents_smoke) %>% 
  mutate_all(as.numeric) %>% 
  na.exclude() %>% 
  select(-pm_mean_sampling_period)

# Use `omcdiag` to assess overall multicollinearity
omcdiag(other_vars, pm_mean_sampling_period)

# Use `imcdiag` to assess multicollinearity between individual variables
imcdiag(other_vars, pm_mean_sampling_period)

# Specify a method to assess
imcdiag(other_vars, pm_mean_sampling_period, method = "Fi")
```

# Initial variables to use for model selection (these have less overlap):
Demographics: income_3level, total_residents_2level, residents_smoke
Home Characteristics: home_type_2level, home_floors_2level, stove_age_3level,
                      chimney_clean_3level, home_mold, home_fireplace
Home Activity: home_act_gas, home_act_smoking
Others (make some categorical): amb_pm_24hr, mean_temp, sums_mean_sampling_day, 
       moisture_closest, stove_grade_3level

# Also try individual chi-square tests and correlations b/w two vars

```{r}
test_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(pm_mean_sampling_day))

chisqu_results <- chisq.test(test_data$home_floors_2level, 
                             test_data$home_windows_2level)
chisqu_results

cor_results <- cor.test(test_data$temp_indoor_max, 
                        test_data$mean_temp)
cor_results
```

If variables have very high overlap, then only use the one with the strongest
association with PM for consideration in the model.

Once variables are narrowed down, use best subsets selection (exposure_models.RMD) 
to help inform a final model.

Run sensitivity analyses to see if adding other variables back in makes a difference.