### 2020-06-24 indoor_temp_rh_cleaning.Rmd

---
title: "KidsAIR indoor temp and humidity cleaning"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 23 June 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
```

# Load, format, and save indoor temp/rh data
```{r}
##### WMT #####
# Load data
temp_rh_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/qtrack.csv"), 
                                  na = c("NULL", ""))

wmt_temp_rh <- temp_rh_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         area = "WMT") %>% 
  # replace -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min) %>% 
  arrange(home_winter_id)

# Save as RDS 
write_rds(wmt_temp_rh, paste0(file_path, "Output/wmt_temp_rh.rds"))

##### NN #####
# Load data
temp_rh_initial_nn <- read_csv(paste0(file_path, "Input/NN/qtrack.csv"), 
                                  na = c("NULL", ""))

nn_temp_rh <- temp_rh_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         area = "NN") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min) %>% 
  arrange(home_winter_id)

# Save as RDS 
write_rds(nn_temp_rh, paste0(file_path, "Output/nn_temp_rh.rds"))

##### AK #####
# Load data
temp_rh_initial_ak <- read_csv(paste0(file_path, "Input/AK/qtrack.csv"), 
                                  na = c("NULL", ""))

ak_temp_rh <- temp_rh_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         area = "AK") %>% 
  # replace -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min) %>% 
  arrange(home_winter_id)

# Save as RDS 
write_rds(ak_temp_rh, paste0(file_path, "Output/ak_temp_rh.rds"))
```

# Join temp/rh data from different areas
```{r}
wmt_temp_rh <- read_rds(paste0(file_path, "Output/wmt_temp_rh.rds"))
nn_temp_rh <- read_rds(paste0(file_path, "Output/nn_temp_rh.rds"))
ak_temp_rh <- read_rds(paste0(file_path, "Output/ak_temp_rh.rds"))

kids_indoor_temp_rh <- rbind(wmt_temp_rh, nn_temp_rh, ak_temp_rh)

write_rds(kids_indoor_temp_rh, paste0(file_path, "Output/kids_indoor_temp_rh.rds"))
```

# Clean up all temp/rh data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_indoor_temp_rh <- read_rds(paste0(file_path, "Output/kids_indoor_temp_rh.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))

indoor_temp_rh_clean <- kids_indoor_temp_rh %>% 
  mutate(area = as.factor(area)) %>% 
  right_join(kids_linked_ids, by = c("area", "home_winter_id"))

summary(indoor_temp_rh_clean)

write_rds(indoor_temp_rh_clean, paste0(file_path, "Output/indoor_temp_rh_clean.rds"))
```