### 2020-08-25 lrti_tracking.Rmd
lrti data dates

---
title: "Kids LRTI tracking"
author: "Ethan Walker"
date: "Started 6 April 2020, Updated 28 July 2020"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

health_exposure_data_sampling_day <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))

check_data <- health_exposure_data_sampling_day %>% 
  filter(child_id_num == 68 & area == "NN")
  filter(home == "CH110")

health_data_merged <- 
  read_rds(paste0(file_path, "Output/health_data_merged.rds"))

wmt_chw <- read_xlsx(paste0(file_path, "input/WMT/childhomewinter.xlsx")) %>% 
  mutate(area = "WMT") %>% 
  rename(paul_comments = 19)
nn_chw <- read_xlsx(paste0(file_path, "input/NN/childhomewinter.xlsx")) %>% 
  mutate(area = "NN") %>% 
  rename(paul_comments = 19)
ak_chw <- read_xlsx(paste0(file_path, "input/AK/childhomewinter.xlsx")) %>% 
  mutate(area = "AK") %>% 
  rename(paul_comments = 19)

child_home_winter <- rbind(wmt_chw, nn_chw, ak_chw) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         paul_lrti = AnyARI) %>% 
  select(area, home_winter_id, child_id_num, paul_lrti, paul_comments)  
  #filter(paul_lrti %in% c("Yes", "Debating"))

demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) %>% 
  filter(child_id_num == 52)
  select(child_id_num, child_id_char, area, winter_id)
```

# Diagnosing LRTI from the data

Is there a specific protocol Paul uses for diagnosing LRTI, or is it case-by-case
based on the given information for each child? We'll need to have some kind of
protocol/algorithm for publications, I would think.

Other questions:
- LRTI and ARI have been used interchangably. Our actual outcome is LRTI, yes?
  - Is there a clinical definition we/Paul have been following when diagnosing cases?
  - LRTI: bronchiolitis, pneumonia, croupe? Case definition up to Paul.
  - Primary diagnosis is based on med recs, but secondary outcomes could be from
    surveys. Also see how well survey data predicts Pauls diagnosis.
- How does Paul initially narrow down which cases to look at? Or does he look
  at each CSQ and med visit form?
- How far has the team made it (date/winter) in diagnosing LRTI? Winter 17/18?
  - Could touch base with Nick, but probably through winter 17/18.
- Has this process been done/started for NN/AK, or just WMT?
  - No. Surveys may be worse for these areas, but med rec data could be better
    through Indian Health Services.
  - At risk period could be difficult for these sites
- Where did the med_visit_ari variable come from?

Notes and specific documentation for steps to identify LRTI cases:
- Double check the filter method of "yes" for sick/still sick
- Ask Kathrene about med rec ARI vs med visit ARI and where they came from
- Was there a variable where parents called us to report sickness?
-

Next steps
- Ask Kathrene some questions from above
  - Have her upload final health data for NN/AK
- Push to get the final data input, QA/QC, uploaded to Box
- Go through winter 18/19 data line by line and flag cases that could be 
  respiratory.
  - Do this for all sites, and start on other sites for all years.
- Also flag other potential secondary outcomes that we could look at.

```{r}
# LRTI Diagnosis data
lrti_diagnosis_data <- health_data_merged %>% 
  # select relevant health data
  select(area, home, home_winter_id, winter_id, child_id_num, 
         visit_date, csq_date:sibling_school, treatment_assigned) %>% 
  filter(!is.na(visit_date)) %>% 
  full_join(child_home_winter, by = c("area", "home_winter_id", "child_id_num")) %>%

# 6 homes have LRTI data, but actually dropped. Notes on those homes below:
# AK HWID 19/Child ID 25: No record that they participated except initial med visits/CSQs. Likely dropped.
# AK HWID 83/Child ID 70: Dropped prior to winter 1
# NN HWID 32/Child ID 45: No record that they participated except initial med visits/CSQs. Likely dropped.
# NN HWID 162/Child ID 171: Dropped prior to winter 1
# NN HWID 164/Child ID 173: Dropped prior to winter 1
# NN HWID 213/Child ID 197: Dropped prior to winter 1
# I confirmed that all of the above homes dropped, so filtering them from dataset

  filter(!is.na(home)) %>% 
  select(area, home, child_id_num, home_winter_id, winter_id, 
         paul_lrti, paul_comments,
         visit_date:treatment_assigned) %>% 
  mutate(lrti_any_winter = if_else(!is.na(paul_lrti), 1, 0)) %>% 
  group_by(home) %>% 
  mutate(lrti_any_winter = sum(lrti_any_winter)) %>% 
  select(area, home, lrti_any_winter, child_id_num, home_winter_id, winter_id, 
         paul_lrti, paul_comments,
         visit_date:treatment_assigned) %>% 
  filter(lrti_any_winter > 0) %>%

# This filtered data contains all observations for kids with LRTI diagnosis of "Yes" or "Debating"
# Next steps will be to:
  # Select necessary variables to identify dates and other info for analysis
  # Identify where they left off assigning dates previously
  # Look through LRTI "No" designations and make sure Paul didn't miss any potential diagnoses
  # Assign inital dates for those that don't have them
    # Make notes along the way for a meeting with Paul to confirm final dates

  mutate(child_id_num = as.numeric(child_id_num),
         home_winter_id = as.numeric(home_winter_id)) %>% 
  arrange(area, child_id_num, home_winter_id)

write_rds(lrti_diagnosis_data, paste0(file_path, "Output/lrti_diagnosis_data.rds"))
write_csv(lrti_diagnosis_data, paste0(file_path, "Output/lrti_diagnosis_data.csv", na = " "))

lrti_notes_decisions <- lrti_diagnosis_data %>% 
  distinct(area, child_id_num, home_winter_id, .keep_all = TRUE) %>% 
  select(area, home, child_id_num, home_winter_id, winter_id, 
         paul_lrti, paul_comments) %>% 
  mutate(notes_initial = "",
         start_date_initial = "",
         end_date_initial = "",
         notes_final = "",
         lrti_final = "",
         lrti_start_date_final = "",
         lrti_end_date_final = "")

write_rds(lrti_notes_decisions, paste0(file_path, "Output/lrti_notes_decisions.rds"))
write_csv(lrti_notes_decisions, paste0(file_path, "Output/lrti_notes_decisions.csv", na = " "))
```

# Merge dataset with final LRTI notes and dates, then assign person-time
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

health_exposure_data_sampling_day <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))

lrti_notes_decisions <- read_xlsx(paste0(file_path, "Input/lrti_notes_decisions.xlsx")) %>% 
  filter(lrti_final == "Yes") %>% 
  select(area, home, child_id_num, home_winter_id, winter_id, lrti_final, 
         lrti_start_date_final, lrti_end_date_final, notes_final, total_cases_this_winter)

wmt_homewinter_initial <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  filter(winterid == 1 | winterid ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = ymd(interventiondate),
         area = "WMT") %>% 
  select(area, home_winter_id, intervention_date)

nn_homewinter_initial <- read_csv(paste0(file_path, "Input/NN/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  filter(winterid == 1 | winterid ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         area = "NN") %>% 
  select(area, home_winter_id, intervention_date)

ak_homewinter_initial <- read_csv(paste0(file_path, "Input/AK/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  filter(winterid == 1 | winterid ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         area = "AK") %>% 
  select(area, home_winter_id, intervention_date)

intervention_dates <- rbind(wmt_homewinter_initial, 
                            nn_homewinter_initial,
                            ak_homewinter_initial)

lrti_yes <- lrti_notes_decisions %>% 
  mutate(child_id_num = as.character(child_id_num),
         home_winter_id = as.character(home_winter_id),
         winter_id = as.character(winter_id)) %>% 
  # join lrti data with full dataset
  full_join(health_exposure_data_sampling_day, 
            by = c("area", "home", "child_id_num", "home_winter_id", "winter_id")) %>% 
  filter(lrti_final == "Yes") %>% 
  # select vars for calculating person-time
  select(area, home, child_id_num, home_winter_id, winter_id, lrti_final, 
         lrti_start_date_final, lrti_end_date_final, notes_final, total_cases_this_winter,
         csq_date) %>% 
  left_join(intervention_dates, by = c("area", "home_winter_id")) %>% 
  group_by(area, child_id_num, winter_id) %>% 
  arrange(area, child_id_num, winter_id, csq_date) %>% 
  # var for person-time start date for each winter/child combo
  # this date is intervention date + 14 days
  mutate(person_time_start = intervention_date + 14) %>% 
  filter(!is.na(csq_date)) %>% 
  # var for person-time end date for each winter/child combo
  # this date is the earliest of the last csq or april 30 of the study year
  mutate(person_time_end = last(csq_date),
         latest_date = "04-30") %>% 
  separate(person_time_end, into = c("pte_year", "pte_mo", "pte_day"), sep = "-", remove = FALSE) %>% 
  unite(latest_date, c("pte_year", "latest_date"), sep = "-") %>% 
  select(-pte_mo, -pte_day) %>% 
  mutate(latest_date = ymd(latest_date),
         person_time_end = if_else(person_time_end > latest_date,
                                   latest_date, person_time_end),
         person_time_possible = person_time_end - person_time_start) %>% 
  # person-time possible is number of days from start to end date for each child/winter combo
  group_by(area, child_id_num, winter_id, lrti_start_date_final) %>% 
  # end of lrti is the lrti end date +14 days
  mutate(lrti_end_date_plus14 = ymd(lrti_end_date_final) + 14,
         lrti_end_date_plus14 = if_else(lrti_end_date_plus14 > latest_date,
                                 latest_date, lrti_end_date_plus14)) %>% 
  # filter out lrti cases that start later than person-time end date (2 cases)
  filter(lrti_start_date_final <= person_time_end) %>% 
  # calculate days with LRTI
  mutate(person_time_lrti = ymd(lrti_end_date_plus14) - ymd(lrti_start_date_final)) %>% 
  distinct(area, child_id_num, winter_id, lrti_start_date_final, .keep_all = TRUE) %>% 
  group_by(area, child_id_num, winter_id) %>% 
  # calculate total LRTI days per child/winter combo
  mutate(person_time_lrti_total = sum(person_time_lrti)) %>% 
  ungroup() %>% 
  #distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  # new var for time without lrti
  mutate(person_time_no_lrti_total = person_time_possible - person_time_lrti_total) %>% 
  arrange(area, child_id_num, winter_id)

lrti_no <- lrti_notes_decisions %>% 
  mutate(child_id_num = as.character(child_id_num),
         home_winter_id = as.character(home_winter_id),
         winter_id = as.character(winter_id)) %>% 
  # join lrti data with full dataset
  full_join(health_exposure_data_sampling_day, 
            by = c("area", "home", "child_id_num", "home_winter_id", "winter_id")) %>% 
  filter(is.na(lrti_final)) %>% 
  # select vars for calculating person-time
  select(area, home, child_id_num, home_winter_id, winter_id, lrti_final, 
         lrti_start_date_final, lrti_end_date_final, notes_final, total_cases_this_winter,
         csq_date) %>% 
  left_join(intervention_dates, by = c("area", "home_winter_id")) %>% 
  group_by(area, child_id_num, winter_id) %>% 
  arrange(area, child_id_num, winter_id, csq_date) %>% 
  # var for person-time start date for each winter/child combo
  # this date is intervention date + 14 days
  mutate(person_time_start = intervention_date + 14) %>% 
  filter(!is.na(csq_date)) %>% 
  # var for person-time end date for each winter/child combo
  # this date is the earliest of the last csq or april 30 of the study year
  mutate(person_time_end = last(csq_date),
         latest_date = "04-30") %>% 
  separate(person_time_end, into = c("pte_year", "pte_mo", "pte_day"), sep = "-", remove = FALSE) %>% 
  unite(latest_date, c("pte_year", "latest_date"), sep = "-") %>% 
  select(-pte_mo, -pte_day) %>% 
  mutate(latest_date = ymd(latest_date),
         person_time_end = if_else(person_time_end > latest_date,
                                   latest_date, person_time_end),
         person_time_end = if_else(person_time_end < person_time_start, last(csq_date), person_time_end),
         person_time_possible = as.numeric(person_time_end - person_time_start),
         person_time_possible = if_else(person_time_possible < 0, 0,
                                        person_time_possible)) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  arrange(area, child_id_num, winter_id)

lrti_final_data <- plyr::rbind.fill(lrti_yes, lrti_no) %>% 
  arrange(area, child_id_num, winter_id)

write_csv(lrti_final_data, paste0(file_path, "Output/lrti_final_data.csv"), na = "")
```