### 2020-10-28 main_analysis_primary_models.Rmd
analysis updates new RMD for secondary analyses

---
title: 'KidsAIR: Main analysis primary models'
author: "Ethan Walker"
date: "Started 27 Oct 2020, Updated 28 Oct 2020"
output:
  html_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r, eval=TRUE, include=TRUE}
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
library(MASS)
library(faraway)
library(influence.ME)
```

```{r}
# Load analysis dataset

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

analysis_data <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) 
  rownames_to_column() %>% 
  #mutate(filter_outliers = if_else(rowname %in% c(57), 1, 0)) %>% 
  mutate(log_person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 2.33, log_person_time_at_risk)) %>% 
  mutate(person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 72, person_time_at_risk)) 
  #filter(lrti_incidence < 0.7) %>% 
  filter(filter_outliers == 0)

health_exposure_data_sampling_day <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
```

Intent-to-treat model framework: LRTI outcome

What's wrong with the QQ plot?
What does the half-normal quantile plot mean?
Does the residual vs fitted plot look appropriate?
```{r}
# Run mixed model and diagnostic plots from above

model_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
                          offset(log_person_time_at_risk), data = analysis_data)

#summary(model_results)

# plot residuals/diagnostics
res <- residuals(model_results, type="deviance")
plot(predict(model_results), res)
abline(h=0, lty=2)
qqnorm(res)
qqline(res)
halfnorm(residuals(model_results))

#influential <- influence(model_results, obs = TRUE)
#cooks <- cooks.distance.estex(influential, sort = TRUE)
#plot.estex(influential, which = "cook")
# Most influential datapoints: 57, 102, 280, 396, 61, 58

# plot results
tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_assigned", "", term)) %>% 
  dplyr::select(term, estimate, p.value, conf.low, conf.high)
#tidy_results

plot_estimates <- tidy_results %>%
  ggplot() +
  geom_point(aes(x=term, y=estimate), size = 4) +
  #scale_shape_manual(values = c(15, 16, 17, 18, 13, 9)) +
  geom_errorbar(aes(x=term, ymin=conf.low, ymax=conf.high), 
                size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  #ggtitle(label = "Systolic BP ITT models") +
  labs(y = "LRTI events per child-week at risk \n compared to placebo treatment") +
  labs(x = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 25,
                                     hjust = .8, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "none") 
plot_estimates
```

Intent-to-treat model framework: PM outcome

What's wrong with the residual vs fitted plot?
This more model is more influenced by additional covariates than the LRTI ITT model
```{r}
# Run mixed model and diagnostic plots from above

model_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + child_age + 
                       (1 | home:cohort:area), data = analysis_data)

#summary(model_results)

# plot residuals/diagnostics
res <- residuals(model_results, type="deviance")
plot(predict(model_results), res)
abline(h=0, lty=2)
qqnorm(res)
qqline(res)
halfnorm(residuals(model_results))

#influential <- influence(model_results, obs = TRUE)
#cooks <- cooks.distance.estex(influential, sort = TRUE)
#plot.estex(influential, which = "cook")

# plot results
tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_assigned", "", term)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high)
#tidy_results

plot_estimates <- tidy_results %>%
  ggplot() +
  geom_point(aes(x=term, y=estimate), size = 4) +
  #scale_shape_manual(values = c(15, 16, 17, 18, 13, 9)) +
  geom_errorbar(aes(x=term, ymin=conf.low, ymax=conf.high), 
                size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  #ggtitle(label = "Systolic BP ITT models") +
  labs(y = "Difference in PM2.5 \n compared to placebo") +
  labs(x = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 25,
                                     hjust = .8, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "none") 
plot_estimates
```

Exposure-response model framework

First assess simple associations between LRTI and covariates
Start with covariates that are associated with PM2.5 from the exposure paper
Run these simple models with Placebo homes to get an idea of associations at "baseline"
Then select a priori confounders based on simple associations and other literature

```{r}
placebo_analysis_data <- analysis_data %>% 
  filter(treatment_assigned == "Placebo")

covar_model <- glmer(lrti_events ~ home_bedrooms + (1 | area), data = placebo_analysis_data,
                   family = poisson(link = "log"), offset = log_person_time_at_risk)
summary(covar_model)

# Selected vars:
# education_3level, residents_smoke
```

```{r}
# Run mixed model and diagnostic plots from above

model_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + child_age + 
                          education_3level + residents_smoke + 
                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
                          data = analysis_data)

#summary(model_results)

# plot residuals/diagnostics
res <- residuals(model_results, type="deviance")
plot(predict(model_results), res)
abline(h=0, lty=2)
qqnorm(res)
qqline(res)
halfnorm(residuals(model_results))

#influential <- influence(model_results, obs = TRUE)
#cooks <- cooks.distance.estex(influential, sort = TRUE)
#plot.estex(influential, which = "cook")
# Most influential data points: 58, 19, 420, 85, 301, 383, 413

# plot results
tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = "PM2.5") %>% 
  dplyr::select(term, estimate, p.value, conf.low, conf.high)
#tidy_results

plot_estimates <- tidy_results %>%
  ggplot() +
  geom_point(aes(x=term, y=estimate), size = 4) +
  #scale_shape_manual(values = c(15, 16, 17, 18, 13, 9)) +
  geom_errorbar(aes(x=term, ymin=conf.low, ymax=conf.high), 
                size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  #ggtitle(label = "Systolic BP ITT models") +
  labs(y = "LRTI cases per IQR increase in PM2.5") +
  labs(x = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "none") 
plot_estimates
```

Logistic regression framework

```{r}
# Run mixed model and diagnostic plots from above

model_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + 
                       (1 | home:cohort:area), data = analysis_data, 
                       family = binomial, offset = log_person_time_at_risk)

#summary(model_results)

# plot residuals/diagnostics
res <- residuals(model_results, type="deviance")
plot(predict(model_results), res)
abline(h=0, lty=2)
qqnorm(res)
qqline(res)
halfnorm(residuals(model_results))

#influential <- influence(model_results, obs = TRUE)
#cooks <- cooks.distance.estex(influential, sort = TRUE)
#plot.estex(influential, which = "cook")

# plot results
tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_assigned", "", term)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high)
#tidy_results

plot_estimates <- tidy_results %>%
  ggplot() +
  geom_point(aes(x=term, y=estimate), size = 4) +
  #scale_shape_manual(values = c(15, 16, 17, 18, 13, 9)) +
  geom_errorbar(aes(x=term, ymin=conf.low, ymax=conf.high), 
                size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  #ggtitle(label = "Systolic BP ITT models") +
  labs(y = "Risk of LRTI event compared to Placebo") +
  labs(x = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "none") 
plot_estimates
```