### 2020-12-24 stove_grade_usage_cleaning.Rmd

---
title: "KidsAIR wood stove grades and usage"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 18 Oct 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(gridExtra)
```

# Load, format, and save wood stove grade data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

##### WMT #####
# Load data
stove_grades_initial_wmt <- read_xlsx(paste0(file_path, "Input/WMT/stove.xlsx"), 
                                  na = c("NULL", ""))

wmt_stove_grades <- stove_grades_initial_wmt %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "WMT") %>% 
  select(home_id, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id) %>% 
  filter(!is.na(stove_grade)) %>% 
  filter(home_id != 0 & home_id != 999)

# Save as RDS
write_rds(wmt_stove_grades, paste0(file_path, "Output/wmt_stove_grades.rds"))

##### NN #####
# Load data
stove_grades_initial_nn <- read_xlsx(paste0(file_path, "Input/NN/stove.xlsx"), 
                                  na = c("NULL", ""))

nn_stove_grades <- stove_grades_initial_nn %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "NN") %>% 
  select(home_id, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id) %>% 
  filter(!is.na(stove_grade)) %>% 
  filter(home_id != 0 & home_id != 999)

# Save as RDS
write_rds(nn_stove_grades, paste0(file_path, "Output/nn_stove_grades.rds"))

##### AK #####
# Load data
stove_grades_initial_ak <- read_xlsx(paste0(file_path, "Input/AK/stove.xlsx"), 
                                  na = c("NULL", ""))

ak_stove_grades <- stove_grades_initial_ak %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "AK") %>% 
  select(home_id, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id) %>% 
  filter(!is.na(stove_grade)) %>% 
  filter(home_id != 0 & home_id != 999)

# Save as RDS
write_rds(ak_stove_grades, paste0(file_path, "Output/ak_stove_grades.rds"))
```

# Join stove grade data from different areas
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

wmt_stove_grades <- read_rds(paste0(file_path, "Output/wmt_stove_grades.rds"))
nn_stove_grades <- read_rds(paste0(file_path, "Output/nn_stove_grades.rds"))
ak_stove_grades <- read_rds(paste0(file_path, "Output/ak_stove_grades.rds"))

kids_stove_grades <- rbind(wmt_stove_grades, nn_stove_grades, ak_stove_grades)

write_rds(kids_stove_grades, paste0(file_path, "Output/kids_stove_grades.rds"))
```

# Clean up all stove grade data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_stove_grades <- read_rds(paste0(file_path, "Output/kids_stove_grades.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))

stove_grades_clean <- kids_stove_grades %>% 
  mutate(area = as.factor(area)) %>% 
  # Mark gave a couple of "D" grades when there was not enough info to score the stove
  # removing those "D" grades from final dataset
  filter(stove_grade != "D") %>% 
  select(-home_id) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
  arrange(area, home, winter_id)

summary(stove_grades_clean)

# Some homes/stoves were graded during both winters, typically when new photos were added
# Since photos were added in these cases, grades may have changed from Winter 1 to Winter 2
# We can't make the assumption that the stove was exactly the same for both winters,
# so it will be best to use the grade corresponding to each winter for analysis.

write_rds(stove_grades_clean, paste0(file_path, "Output/stove_grades_clean.rds"))

# Dataset showing homes with no stove grade
stove_grades_na <- stove_grades_clean %>% 
  filter(is.na(stove_grade)) %>% 
  distinct(home_id) 
```

###############################################

# Load, format, and save wood stove use data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

##### WMT #####
# Load data
stove_use_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/woodstoveusage.csv"), 
                                  na = c("NULL", ""))

wmt_stove_use <- stove_use_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         wood_collect = as.factor(WhenGet),
         wood_collect_method = as.factor(HowGet),
         wood_cords = as.numeric(Cords),
         area = "WMT") %>% 
  select(home_winter_id, area, burn_level, wood_type, wood_type_other,
         wood_collect, wood_collect_method, wood_cords) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
write_rds(wmt_stove_use, paste0(file_path, "Output/wmt_stove_use.rds"))

##### NN #####
# Load data
stove_use_initial_nn <- read_csv(paste0(file_path, "Input/NN/woodstoveusage.csv"), 
                                  na = c("NULL", ""))

nn_stove_use <- stove_use_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         wood_collect = as.factor(WhenGet),
         wood_collect_method = as.factor(HowGet),
         wood_cords = as.numeric(Cords),
         area = "NN") %>% 
  select(home_winter_id, area, burn_level, wood_type, wood_type_other,
         wood_collect, wood_collect_method, wood_cords) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
write_rds(nn_stove_use, paste0(file_path, "Output/nn_stove_use.rds"))

##### AK #####
# Load data
stove_use_initial_ak <- read_csv(paste0(file_path, "Input/AK/woodstoveusage.csv"), 
                                  na = c("NULL", ""))

ak_stove_use <- stove_use_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         wood_collect = as.factor(WhenGet),
         wood_collect_method = as.factor(HowGet),
         wood_cords = as.numeric(Cords),
         area = "AK") %>% 
  select(home_winter_id, area, burn_level, wood_type, wood_type_other,
         wood_collect, wood_collect_method, wood_cords) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
write_rds(ak_stove_use, paste0(file_path, "Output/ak_stove_use.rds"))
```

# Join stove use data from different areas
```{r}
wmt_stove_use <- read_rds(paste0(file_path, "Output/wmt_stove_use.rds"))
nn_stove_use <- read_rds(paste0(file_path, "Output/nn_stove_use.rds"))
ak_stove_use <- read_rds(paste0(file_path, "Output/ak_stove_use.rds"))

kids_stove_use <- rbind(wmt_stove_use, nn_stove_use, ak_stove_use)

write_rds(kids_stove_use, paste0(file_path, "Output/kids_stove_use.rds"))
```

# Clean up all stove use data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_stove_use <- read_rds(paste0(file_path, "Output/kids_stove_use.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))

stove_use_clean <- kids_stove_use %>% 
  mutate(area = as.factor(area)) %>% 
  replace_with_na(replace = list(wood_cords = -999)) %>%
  replace_with_na(replace = list(wood_cords = 0)) %>%
  mutate(wood_collect = factor(wood_collect,
                               levels = c("Less than 1 week",
                                          "Between 1 week and 1 month",
                                          "Between 1 month and 3 months",
                                          "Between 3 months and 6 months",
                                          "Between 6 months and 1 year",
                                          "More than 1 year"),
                               labels = c("< 1 week",
                                          "1 week to 1 mo",
                                          "1 mo to 3 mo",
                                          "3 mo to 6 mo",
                                          "6 mo to 1 year",
                                          "> 1 yr"))) %>% 
  right_join(kids_linked_ids, by = c("area", "home_winter_id"))

summary(stove_use_clean)

write_rds(stove_use_clean, paste0(file_path, "Output/stove_use_clean.rds"))
```

#####################################

# Load, format, and KW monitor data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

##### WMT #####
# Load data
kw_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/filtretechange.csv"), 
                                  na = c("NULL", "", "...", -99, -999, -9999))

kw_wmt <- kw_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         area = "WMT") %>% 
  select(home_winter_id, area, kw_date, kw_hours) %>% 
  filter(!is.na(kw_hours)) %>% 
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_wmt, paste0(file_path, "Output/kw_wmt.rds"))

##### NN #####
# Load data
kw_initial_nn <- read_csv(paste0(file_path, "Input/NN/filtretechange.csv"), 
                                  na = c("NULL", "", "...", -99, -999, -9999))

kw_nn <- kw_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         area = "NN") %>% 
  select(home_winter_id, area, kw_date, kw_hours) %>% 
  filter(!is.na(kw_hours)) %>%  
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_nn, paste0(file_path, "Output/kw_nn.rds"))

##### AK #####
# Load data
kw_initial_ak <- read_csv(paste0(file_path, "Input/AK/filtretechange.csv"), 
                                  na = c("NULL", "", "...", -99, -999, -9999))

kw_ak <- kw_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         area = "AK") %>% 
  select(home_winter_id, area, kw_date, kw_hours) %>%  
  filter(!is.na(kw_hours)) %>% 
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_ak, paste0(file_path, "Output/kw_ak.rds"))
```

# Join KW hour data from different areas
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kw_wmt <- read_rds(paste0(file_path, "Output/kw_wmt.rds"))
kw_nn <- read_rds(paste0(file_path, "Output/kw_nn.rds"))
kw_ak <- read_rds(paste0(file_path, "Output/kw_ak.rds"))

kids_kw <- rbind(kw_wmt, kw_nn, kw_ak) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, kw_date)

write_rds(kids_kw, paste0(file_path, "Output/kids_kw.rds"))
```

# Clean up all KW hour data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_kw <- read_rds(paste0(file_path, "Output/kids_kw.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
treatments_masked <- read_csv(paste0(file_path, "Input/KidsAIRrnd_masked.csv"), na = "NULL") %>% 
  mutate(home = gsub(" ", "", x = home),
         home = gsub("_", "", x = home))

kids_kw_clean <- kids_kw %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%  
  filter(!is.na(home)) %>% 
  arrange(area, home_winter_id, kw_date) %>% 
  group_by(area, home_winter_id) %>% 
  mutate(kw_hours_last = last(kw_hours),
         diff_check = if_else(kw_hours_last < kw_hours, 1, 0),
         diff_check2 = kw_hours_last - kw_hours)  
  distinct(area, home_winter_id, diff_check, .keep_all = TRUE)

summary(kids_kw_clean)

kw_summary <- kids_kw_clean %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>% 
  distinct(area, home, kw_hours_total) %>% 
  ungroup() %>% 
  filter(area == "WMT") %>% 
  #group_by(newtx) %>% 
  summarize(mean(kw_hours_total),
            sd(kw_hours_total),
            n(),
            min(kw_hours_total),
            median(kw_hours_total),
            max(kw_hours_total))
kw_summary

# Some homes/stoves were graded during both winters, typically when new photos were added
# Since photos were added in these cases, grades may have changed from Winter 1 to Winter 2
# We can't make the assumption that the stove was exactly the same for both winters,
# so it will be best to use the grade corresponding to each winter for analysis.

write_rds(kids_kw_clean, paste0(file_path, "Output/kids_kw_clean.rds"))
```

#####################################

# Load, format, and KW compliance report
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

##### WMT #####
# Load data
kw_report_wmt <- read_csv(paste0(file_path, "Input/WMT/filtrete_report.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         home = as.factor(Home),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         kw_notes = as.character(Notes),
         hours_possible = as.numeric(Hours),
         kw_expected = as.numeric(Expected),
         kw_perc_expected = as.numeric(PercExp),
         filter_type = as.character(FType),
         area = "WMT") %>% 
  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
         hours_possible, kw_expected, kw_perc_expected) %>% 
  #filter(!is.na(kw_hours)) %>% 
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_report_wmt, paste0(file_path, "Output/kw_report_wmt.rds"))

##### NN #####
# Load data
kw_report_nn <- read_csv(paste0(file_path, "Input/NN/filtrete_report.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         home = as.factor(Home),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         kw_notes = as.character(Notes),
         hours_possible = as.numeric(Hours),
         kw_expected = as.numeric(Expected),
         kw_perc_expected = as.numeric(PercExp),
         filter_type = as.character(FType),
         area = "NN") %>% 
  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
         hours_possible, kw_expected, kw_perc_expected) %>% 
  #filter(!is.na(kw_hours)) %>% 
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_report_nn, paste0(file_path, "Output/kw_report_nn.rds"))

##### AK #####
# Load data
kw_report_ak <- read_csv(paste0(file_path, "Input/AK/filtrete_report.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         home = as.factor(Home),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         kw_notes = as.character(Notes),
         hours_possible = as.numeric(Hours),
         kw_expected = as.numeric(Expected),
         kw_perc_expected = as.numeric(PercExp),
         filter_type = as.character(FType),
         area = "AK") %>% 
  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
         hours_possible, kw_expected, kw_perc_expected) %>% 
  #filter(!is.na(kw_hours)) %>% 
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_report_ak, paste0(file_path, "Output/kw_report_ak.rds"))
```

# Join KW report data from different areas
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kw_report_wmt <- read_rds(paste0(file_path, "Output/kw_report_wmt.rds"))
kw_report_nn <- read_rds(paste0(file_path, "Output/kw_report_nn.rds"))
kw_report_ak <- read_rds(paste0(file_path, "Output/kw_report_ak.rds"))

kids_kw_report <- rbind(kw_report_wmt, kw_report_nn, kw_report_ak) %>% 
  dplyr::select(-home) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, kw_date)

write_rds(kids_kw_report, paste0(file_path, "Output/kids_kw_report.rds"))
```

# Clean up all KW report data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_kw_report <- read_rds(paste0(file_path, "Output/kids_kw_report.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
treatments_masked <- read_csv(paste0(file_path, "Input/KidsAIRrnd_masked.csv"), na = "NULL") %>% 
  mutate(home = gsub(" ", "", x = home),
         home = gsub("_", "", x = home))

kids_kw_report_clean <- kids_kw_report %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%  
  #filter(winter_id == 1) %>% 
  arrange(area, home_winter_id) %>% 
  mutate(filter_type = fct_collapse(filter_type,
                                    "Honeywell" = "Honeywell",
                                    "Large Filtrete" = "Large Filtrete",
                                    "Small Filtrete" = "Small Filtrete",
                                    "Winix" = c("Winix", "winix"))) %>% 
  mutate(kw_perc_expected = if_else(kw_hours == -999, 999999, kw_perc_expected),
         kw_perc_expected = if_else(kw_hours == 1000, 999999, kw_perc_expected),
         kw_perc_expected = abs(kw_perc_expected)) %>% 
  replace_with_na(replace = list(kw_perc_expected = 999999)) %>% 
  mutate(kw_perc_expected = as.numeric(kw_perc_expected)) 
  filter(treatment_assigned == "Filter")

#summary(kids_kw_report_clean)

kw_perc_exp_summary <- kids_kw_report_clean %>% 
  filter(!is.na(kw_perc_expected)) %>% 
  #filter(area == "WMT") %>% 
  group_by(treatment_assigned) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_expected),
            "SD KW % Exp" = sd(kw_perc_expected),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_expected),
            "25Q KW % Exp" = quantile(kw_perc_expected, 0.25),
            "Med KW % Exp" = median(kw_perc_expected),
            "75Q KW % Exp" = quantile(kw_perc_expected, 0.75),
            "Max KW % Exp" = max(kw_perc_expected))
kw_perc_exp_summary

png("Kids_KW_perc_exp.png", width=960, height=480, bg = "white")
grid.table(kw_perc_exp_summary)
dev.off()

boxplot(log(kw_perc_expected) ~ filter_type, data = kids_kw_report_clean)

boxplot(kw_perc_expected ~ filter_type, data = kids_kw_report_clean)

kw_exp_summary <- kids_kw_report_clean %>% 
  filter(!is.na(kw_expected)) %>% 
  #filter(area == "WMT") %>% 
  group_by(filter_type) %>% 
  summarize("Mean KW Exp" = mean(kw_expected),
            "SD KW Exp" = sd(kw_expected),
            "N KW Exp" = n(),
            "Min KW Exp" = min(kw_expected),
            "25Q KW Exp" = quantile(kw_expected, 0.25),
            "Med KW Exp" = median(kw_expected),
            "75Q KW Exp" = quantile(kw_expected, 0.75),
            "Max KW Exp" = max(kw_expected))
kw_exp_summary

fivenum(kids_kw_report_clean$kw_perc_expected)

write_rds(kids_kw_report_clean, paste0(file_path, "Output/kids_kw_report_clean.rds"))
```

# WMT filter unit compliance phone calls conducted by Carolyn
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

wmt_filter_checks <- read_xlsx(paste0(file_path, "Input/WMT/EdChecks-WithComments.xlsx"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, date, 
         filterrunning, 'filter details') %>% 
  mutate(home_winter_id = as.character(homewinterid),
         date = ymd(date),
         filter_level = as.factor(filterrunning),
         filter_comment = as.character(`filter details`),
         area = "WMT") %>% 
  select(home_winter_id, date, 
         filter_level, filter_comment, area) %>% 
  arrange(home_winter_id, date)

write_rds(wmt_filter_checks, paste0(file_path, "Output/wmt_filter_checks.rds"))
```

# Summarize KWhour and PM data; print table
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
pm_data <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
  arrange(area, home, pm_mean_sampling_period) %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  dplyr::select(area, home, pm_mean_sampling_period, treatment_assigned, home_sqft_2level)
kids_kw_report_clean <- 
  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
  dplyr::select(-treatment_assigned)
wmt_filter_checks <- 
  read_rds(paste0(file_path, "Output/wmt_filter_checks.rds"))

kw_perc_exp_summary <- kids_kw_report_clean %>% 
  left_join(pm_data, by = c("area", "home")) %>% 
  right_join(wmt_filter_checks, by = c("area", "home_winter_id")) %>% 
  #filter(!is.na(kw_perc_expected)) %>% 
  #filter(treatment_assigned == "Filter") %>% 
  #distinct(area, home, .keep_all = TRUE) %>% 
  #filter(!is.na(filter_type)) %>% 
  #group_by(area, treatment_assigned, filter_type, home_sqft_2level) %>% 
  filter(area == "WMT") %>% 
  group_by(filter_type, treatment_assigned, filter_level) %>% 
  summarize("N KW % Exp" = n(),
            "Mean KW % Exp" = mean(kw_perc_expected, na.rm = TRUE),
            #"SD KW % Exp" = sd(kw_perc_expected, na.rm = TRUE),
            #"Min KW % Exp" = min(kw_perc_expected, na.rm = TRUE),
            "Med KW % Exp" = median(kw_perc_expected, na.rm = TRUE),
            #"Max KW % Exp" = max(kw_perc_expected),
            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE))
kw_perc_exp_summary

# wmt_filte_checks data seems to validate the KW % expected as compliance data,
# at least for WMT. KW % Exp is lower in homes that reported filter as too low or off.

png("Kids_KW_perc_exp.png", width=1200, height=480, bg = "white")
grid.table(kw_perc_exp_summary)
dev.off()
```

# Extract sampling dates and use KWhour data from PM sampling period
```{r}
wmt_sampling_dates <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = ymd(interventiondate),
         sampling_date = ymd(samplingdate),
         pickup_date = ymd(pickupdate),
         area = "WMT") %>% 
  select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

nn_sampling_dates <- read_csv(paste0(file_path, "Input/NN/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         sampling_date = mdy(samplingdate),
         pickup_date = mdy(pickupdate),
         area = "NN") %>% 
  select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

ak_sampling_dates <- read_csv(paste0(file_path, "Input/AK/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         sampling_date = mdy(samplingdate),
         pickup_date = mdy(pickupdate),
         area = "AK") %>% 
  select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

kids_sampling_dates <- rbind(wmt_sampling_dates, nn_sampling_dates, ak_sampling_dates)

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_kw_clean <- 
  read_rds(paste0(file_path, "Output/kids_kw_clean.rds"))
filter_units <- 
  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
  select(home_winter_id, area, filter_type)

kids_kw_dates <- kids_kw_clean %>% 
  full_join(kids_sampling_dates, by = c("area", "home_winter_id")) %>% 
  mutate(filter_var = if_else(kw_date == sampling_date | kw_date == pickup_date, 1, 0)) %>% 
  filter(filter_var == 1) %>% 
  arrange(area, home, kw_date) %>% 
  mutate(kw_diff = lead(kw_hours) - kw_hours) %>% 
  filter(!is.na(kw_diff)) %>% 
  mutate(date_diff = as.duration(interval(sampling_date, pickup_date)),
         date_diff = as.numeric(date_diff)/86400) %>% 
  left_join(filter_units, by = c("area", "home_winter_id")) %>% 
  mutate(multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Placebo",
                              2.04, 0),
         multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Filter",
                              4.224, multiplier),
         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Placebo",
                              0.888, multiplier),
         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Filter",
                              1.896, multiplier),
         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Placebo",
                              1.896, multiplier),
         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Filter",
                              3.48, multiplier),
         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Placebo",
                              0.0144, multiplier),
         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Filter",
                              0.48, multiplier)) %>% 
  mutate(kw_expected = date_diff*multiplier,
         kw_perc_exp = kw_diff/kw_expected*100)

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
pm_data <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
  arrange(area, home, pm_mean_sampling_period) %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  select(area, home, pm_mean_sampling_period)

kw_perc_exp_summary <- kids_kw_dates %>% 
  left_join(pm_data, by = c("area", "home")) %>% 
  filter(!is.na(kw_perc_exp)) %>% 
  #filter(area == "WMT") %>% 
  group_by(filter_type, treatment_assigned) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_exp),
            "SD KW % Exp" = sd(kw_perc_exp),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_exp),
            "Med KW % Exp" = median(kw_perc_exp),
            "Max KW % Exp" = max(kw_perc_exp),
            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE),
            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
kw_perc_exp_summary

png("Kids_KW_perc_exp_sampling.png", width=1200, height=480, bg = "white")
grid.table(kw_perc_exp_summary)
dev.off()
```

# Use KWhour data only if KWhours increased in subsequent dates
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_kw_clean <- 
  read_rds(paste0(file_path, "Output/kids_kw_clean.rds"))

filter_units <- 
  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
  dplyr::select(home_winter_id, area, filter_type)

wmt_sampling_dates <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  dplyr::select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = ymd(interventiondate),
         sampling_date = ymd(samplingdate),
         pickup_date = ymd(pickupdate),
         area = "WMT") %>% 
  dplyr::select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

nn_sampling_dates <- read_csv(paste0(file_path, "Input/NN/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  dplyr::select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         sampling_date = mdy(samplingdate),
         pickup_date = mdy(pickupdate),
         area = "NN") %>% 
  dplyr::select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

ak_sampling_dates <- read_csv(paste0(file_path, "Input/AK/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  dplyr::select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         sampling_date = mdy(samplingdate),
         pickup_date = mdy(pickupdate),
         area = "AK") %>% 
  dplyr::select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

kids_sampling_dates <- rbind(wmt_sampling_dates, nn_sampling_dates, ak_sampling_dates)

kids_filter_compliance <- kids_kw_clean %>% 
  group_by(area, home_winter_id) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  mutate(max_kw_hours = if_else(kw_hours == max(kw_hours), 1, 0)) %>% 
  filter(max_kw_hours == 1) %>% 
  left_join(kids_sampling_dates, by = c("area", "home_winter_id")) %>% 
  mutate(date_diff = as.duration(interval(intervention_date, kw_date)),
         date_diff = as.numeric(date_diff)/86400) %>% 
  left_join(filter_units, by = c("area", "home_winter_id")) %>% 
  mutate(multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Placebo",
                              2.04, 0),
         multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Filter",
                              4.224, multiplier),
         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Placebo",
                              0.888, multiplier),
         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Filter",
                              1.896, multiplier),
         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Placebo",
                              1.896, multiplier),
         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Filter",
                              3.48, multiplier),
         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Placebo",
                              0.0144, multiplier),
         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Filter",
                              0.48, multiplier)) %>% 
  mutate(kw_expected = date_diff*multiplier,
         kw_perc_exp = kw_hours/kw_expected*100) %>% 
  dplyr::select(area, home_winter_id, winter_id, home, treatment_assigned, 
                kw_date, kw_hours, date_diff, filter_type, multiplier,
                kw_expected, kw_perc_exp)

compliance_summary1 <- kids_filter_compliance %>% 
  group_by(home) %>% 
  mutate(kw_perc_exp_total = mean(kw_perc_exp, na.rm = TRUE)) %>%  
  filter(treatment_assigned == "Filter") %>% 
  filter(!is.na(kw_perc_exp_total)) %>% 
  group_by(treatment_assigned) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_exp_total),
            "SD KW % Exp" = sd(kw_perc_exp_total),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_exp_total),
            "25Q KW % Exp" = quantile(kw_perc_exp_total, 0.25),
            "Med KW % Exp" = median(kw_perc_exp_total),
            "75Q KW % Exp" = quantile(kw_perc_exp_total, 0.75),
            "Max KW % Exp" = max(kw_perc_exp_total))
compliance_summary1

kids_filter_compliance_total <- kids_filter_compliance %>% 
  group_by(home) %>% 
  mutate(kw_perc_exp_total = mean(kw_perc_exp), 
         treatment_assigned = as.character(treatment_assigned),
         perc_exp_quart = cut(kw_perc_exp_total,
                                  breaks = c(0, 30, 64, 100, 1000),
                                  labels = c("0-25 compliance", "25-50 compliance",
                                             "50-75 compliance", "75-100 compliance")),
         perc_exp_quart = as.character(perc_exp_quart),
         filter_compliance = if_else(treatment_assigned == "Filter",
                                     perc_exp_quart, 
                                     treatment_assigned)) %>% 
  distinct(area, home, .keep_all = TRUE)

compliance_summary2 <- kids_filter_compliance %>% 
  group_by(area, home_winter_id) %>% 
  mutate(kw_perc_exp_winter = mean(kw_perc_exp, na.rm = TRUE)) %>%  
  filter(treatment_assigned == "Filter") %>% 
  filter(!is.na(kw_perc_exp_winter)) %>% 
  group_by(winter_id, treatment_assigned) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_exp_winter),
            "SD KW % Exp" = sd(kw_perc_exp_winter),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_exp_winter),
            "25Q KW % Exp" = quantile(kw_perc_exp_winter, 0.25),
            "Med KW % Exp" = median(kw_perc_exp_winter),
            "75Q KW % Exp" = quantile(kw_perc_exp_winter, 0.75),
            "Max KW % Exp" = max(kw_perc_exp_winter))
compliance_summary2

kids_filter_compliance_winter1 <- kids_filter_compliance %>% 
  filter(winter_id == 1) %>% 
  mutate(treatment_assigned = as.character(treatment_assigned),
         perc_exp_quart = cut(kw_perc_exp,
                                  breaks = c(0, 32, 60, 116, 1000),
                                  labels = c("0-25 compliance", "25-50 compliance",
                                             "50-75 compliance", "75-100 compliance")),
         perc_exp_quart = as.character(perc_exp_quart),
         filter_compliance = if_else(treatment_assigned == "Filter",
                                     perc_exp_quart, 
                                     treatment_assigned))

write_rds(kids_filter_compliance_total, paste0(file_path, "Output/kids_filter_compliance_total.rds"))
write_rds(kids_filter_compliance_winter1, paste0(file_path, "Output/kids_filter_compliance_winter1.rds"))

pm_data <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
  arrange(area, home, pm_mean_sampling_period) %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  select(area, home, pm_mean_sampling_period)

kw_perc_exp_summary <- kids_kw_filtered %>% 
  left_join(pm_data, by = c("area", "home")) %>% 
  filter(!is.na(kw_perc_exp)) %>% 
  #filter(area == "WMT") %>% 
  group_by(filter_type, treatment_assigned) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_exp),
            "SD KW % Exp" = sd(kw_perc_exp),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_exp),
            "Med KW % Exp" = median(kw_perc_exp),
            "Max KW % Exp" = max(kw_perc_exp),
            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE),
            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
kw_perc_exp_summary

png("Kids_KW_perc_exp_new.png", width=1200, height=480, bg = "white")
grid.table(kw_perc_exp_summary)
dev.off()  
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_kw_clean <- 
  read_rds(paste0(file_path, "Output/kids_kw_clean.rds"))

filter_units <- 
  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
  select(home_winter_id, area, filter_type)

wmt_sampling_dates <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = ymd(interventiondate),
         sampling_date = ymd(samplingdate),
         pickup_date = ymd(pickupdate),
         area = "WMT") %>% 
  select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

nn_sampling_dates <- read_csv(paste0(file_path, "Input/NN/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         sampling_date = mdy(samplingdate),
         pickup_date = mdy(pickupdate),
         area = "NN") %>% 
  select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

ak_sampling_dates <- read_csv(paste0(file_path, "Input/AK/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         sampling_date = mdy(samplingdate),
         pickup_date = mdy(pickupdate),
         area = "AK") %>% 
  select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

kids_sampling_dates <- rbind(wmt_sampling_dates, nn_sampling_dates, ak_sampling_dates)

kids_kw_filtered <- kids_kw_clean %>% 
  left_join(kids_sampling_dates, by = c("area", "home_winter_id")) %>% 
  mutate(date_diff = as.duration(interval(intervention_date, kw_date)),
         date_diff = as.numeric(date_diff)/86400) %>% 
  left_join(filter_units, by = c("area", "home_winter_id")) %>% 
  distinct(area, home_winter_id, .keep_all = TRUE) %>% 
  mutate(multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Placebo",
                              2.04, 0),
         multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Filter",
                              4.224, multiplier),
         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Placebo",
                              0.888, multiplier),
         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Filter",
                              1.896, multiplier),
         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Placebo",
                              1.896, multiplier),
         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Filter",
                              3.48, multiplier),
         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Placebo",
                              0.0144, multiplier),
         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Filter",
                              0.48, multiplier)) %>% 
  mutate(kw_expected = date_diff*multiplier,
         kw_perc_exp = kw_hours/kw_expected*100)

pm_data <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
  arrange(area, home, pm_mean_sampling_period) %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  select(area, home, pm_mean_sampling_period)

kw_perc_exp_summary <- kids_kw_filtered %>% 
  left_join(pm_data, by = c("area", "home")) %>% 
  filter(!is.na(kw_perc_exp)) %>% 
  filter(kw_perc_exp < 30000) %>% 
  #filter(area == "WMT") %>% 
  group_by(filter_type, treatment_assigned) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_exp, na.rm = TRUE),
            "SD KW % Exp" = sd(kw_perc_exp, na.rm = TRUE),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_exp, na.rm = TRUE),
            "Med KW % Exp" = median(kw_perc_exp, na.rm = TRUE),
            "Max KW % Exp" = max(kw_perc_exp, na.rm = TRUE),
            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE),
            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
kw_perc_exp_summary

# Changing the placebo multipliers above to the same as the filter multiplier
# changes the picture quite a bit. It seems likely that participants ran the 
# placebo filters on high more often than expected, resulting in high % Exp KWhours

png("Kids_KW_perc_exp_new.png", width=1200, height=480, bg = "white")
grid.table(kw_perc_exp_summary)
dev.off()  
```