### 2021-01-21 intervention_questionnaire_summary.Rmd

---
title: "KidsAIR intervention questionnaires summary"
author: "Ethan Walker"
date: "Started 20 Jan 2021, Updated 20 Jan 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_pwq_filter_clean <- read_rds(paste0(file_path, "Output/kids_pwq_filter_clean.rds")) 
kids_pwq_education_clean <- read_rds(paste0(file_path, "Output/kids_pwq_education_clean.rds")) 
kids_kab_filter_clean <- read_rds(paste0(file_path, "Output/kids_kab_filter_clean.rds")) 
kids_kab_education_clean <- read_rds(paste0(file_path, "Output/kids_kab_education_clean.rds")) 
```

# Summarize questionnaires for K99/R00 proposal
```{r}
kids_pwq_filter_summary <- kids_pwq_filter_clean %>% 
  filter(area == "WMT") %>% 
  filter(winter_id == 2) 
summary(kids_pwq_filter_summary)

kids_pwq_education_summary <- kids_pwq_education_clean %>% 
  filter(area == "WMT") %>% 
  filter(winter_id == 2) 
summary(kids_pwq_education_summary)

kids_kab_filter_summary <- kids_kab_filter_clean %>% 
  filter(area == "WMT") %>% 
  filter(winter_id == 2) 
summary(kids_kab_filter_summary)

kids_kab_education_summary <- kids_kab_education_clean %>% 
  filter(area == "WMT") %>% 
  filter(winter_id == 2) 
summary(kids_kab_education_summary)
```