### 2021-06-09 intervention_questionnaire_summary.Rmd
PM analysis updates. Manuscript revisions from co-author comments.

---
title: "KidsAIR intervention questionnaires summary"
author: "Ethan Walker"
date: "Started 20 Jan 2021, Updated 20 Jan 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_pwq_filter_clean <- read_rds(paste0(file_path, "Output/kids_pwq_filter_clean.rds")) 
kids_pwq_education_clean <- read_rds(paste0(file_path, "Output/kids_pwq_education_clean.rds")) 
kids_kab_filter_clean <- read_rds(paste0(file_path, "Output/kids_kab_filter_clean.rds")) 
kids_kab_education_clean <- read_rds(paste0(file_path, "Output/kids_kab_education_clean.rds")) 
```

# Summarize questionnaires for K99/R00 proposal
```{r}
kids_pwq_filter_summary <- kids_pwq_filter_clean %>% 
  filter(treatment_assigned == "Filter") %>% 
  select(area, winter_id, fuse, helpful, continuously) %>% 
  group_by(winter_id) %>% 
  mutate(winter_n = n()) %>% 
  group_by(winter_id, winter_n, helpful) %>% 
  count() %>% 
  mutate(var_perc = round(n/winter_n*100, digits = 1)) %>% 
  ungroup() %>% 
  select(winter_id, helpful, n, var_perc)
kids_pwq_filter_summary

kids_pwq_education_summary <- kids_pwq_education_clean %>% 
  select(area, winter_id, mmuse, mmhelpfulness,
         thuse, thhelpfulness, fsuse, fshelpfulness) %>% 
  group_by(winter_id) %>% 
  mutate(winter_n = n()) %>% 
  group_by(winter_id, winter_n, fshelpfulness) %>% 
  count() %>% 
  mutate(var_perc = round(n/winter_n*100, digits = 1)) %>% 
  ungroup() %>% 
  select(winter_id, fshelpfulness, n, var_perc)
kids_pwq_education_summary

kids_kab_filter_summary <- kids_kab_filter_clean %>% 
  filter(area == "WMT") %>% 
  filter(winter_id == 2) 
summary(kids_kab_filter_summary)

kids_kab_education_summary <- kids_kab_education_clean %>% 
  filter(area == "WMT") %>% 
  filter(winter_id == 2) 
summary(kids_kab_education_summary)
```